---
title: "Data Import & Cleaning (study1adults)"
author: "Adam Stone, PhD" 
date: '`r format(Sys.Date(), "%m-%d-%Y")`'
#date: "September 4, 2017"
output:
  github_document:
    toc: yes
    toc_depth: 2
  html_notebook:
    code_folding: hide
    theme: paper
    highlight: tango
    toc: yes
    toc_depth: 2
    toc_float: yes
    df_print: paged
---

# Importing and Reshaping Data
Here we're going to import the data, remove dropped participants, and reshape the data so story and direction are grouping variables (and the dataset will be more tall than wide). Let's see ALL of our data first (scroll horizontally).

```{r, echo=FALSE, message=FALSE, warning=FALSE}

# Import packages we'll need.
library(tidyverse)
library(stringr)
library(lme4)
library(lmerTest)
library(prettydoc)
library(broom)
library(knitr)
library(xtable)
library(kableExtra)
options(knitr.table.format = "html") 
# Import the data
data <- read_csv('finaladultdata.csv', col_types=
                   cols(
                     id = col_integer(),
                     participant = col_character(),
                     hearing = col_character(),
                     videogroup = col_character(),
                     aoagroup = col_character(),
                     languagegroup = col_character(),
                     maingroup = col_character(),
                     selfrate = col_double(),
                     age = col_double(),
                     signyrs = col_double(),
                     aoasl = col_integer(),
                     acc.fw1 = col_double(),
                     acc.rv2 = col_double(),
                     acc.fw3 = col_double(),
                     acc.rv4 = col_double(),
                     forehead.fw1 = col_double(),
                     forehead.fw3 = col_double(),
                     forehead.rv2 = col_double(),
                     forehead.rv4 = col_double(),
                     eyes.fw1 = col_double(),
                     eyes.fw3 = col_double(),
                     eyes.rv2 = col_double(),
                     eyes.rv4 = col_double(),
                     mouth.fw1 = col_double(),
                     mouth.fw3 = col_double(),
                     mouth.rv2 = col_double(),
                     mouth.rv4 = col_double(),
                     chin.fw1 = col_double(),
                     chin.fw3 = col_double(),
                     chin.rv2 = col_double(),
                     chin.rv4 = col_double(),
                     upperchest.fw1 = col_double(),
                     upperchest.fw3 = col_double(),
                     upperchest.rv2 = col_double(),
                     upperchest.rv4 = col_double(),
                     midchest.fw1 = col_double(),
                     midchest.fw3 = col_double(),
                     midchest.rv2 = col_double(),
                     midchest.rv4 = col_double(),
                     lowerchest.fw1 = col_double(),
                     lowerchest.fw3 = col_double(),
                     lowerchest.rv2 = col_double(),
                     lowerchest.rv4 = col_double(),
                     belly.fw1 = col_double(),
                     belly.fw3 = col_double(),
                     belly.rv2 = col_double(),
                     belly.rv4 = col_double(),
                     left.fw1 = col_double(),
                     left.fw3 = col_double(),
                     left.rv2 = col_double(),
                     left.rv4 = col_double(),
                     right.fw1 = col_double(),
                     right.fw3 = col_double(),
                     right.rv2 = col_double(),
                     right.rv4 = col_double()
                     )
)
#kable(data) %>% kable_styling(bootstrap_options = c("striped", "hover","condensed","responsive"), font_size=7, full_width = F, position = "left")
select(data,-participant)
```

Remove dropped participants (they have no main group name, or no data). These are who we dropped. 

```{r, message=FALSE, warning=FALSE}
nodata <- filter(data, is.na(maingroup)==TRUE)
manual <- filter(data,participant=="Lucinda" | participant=="Joe")
dropped <- rbind(nodata,manual) %>% arrange(id)
data <- data %>%
  anti_join(dropped)
#kable(select(dropped,participant,maingroup,age)) %>% kable_styling(bootstrap_options = c("striped", "hover","condensed"), full_width = F, position = "left")
select(dropped,-participant)
```

Now we'll reshape the data. Based on Rain's UNM talk, this is what Group 1 & 2 saw:
**Note, I've switched the groups around from Rain's UNM talk** 

| No. | Group1 | Group2 |
|:------|:-----|:------|
| 1     | Goldilocks Fwd    | Red Riding Hood Fwd   |
| 2     | Cinderella Rev    | King Midas Rev   |
| 3    | King Midas Fwd    | Cinderella Fwd   |
| 4    | Red Riding Hood Rev    | Goldilocks Rev   |

Let's add that information to our data when we reshape it. You can look at the code below if you want. 

```{r, message=FALSE, warning=FALSE}
# I tried writing a function to do this using column names as arguments. 
# But after hours of googling I couldn't figure it out! So just copying/pasting code here.
data.acc <- data %>%
  select(id,acc.fw1:acc.rv4) %>%
  gather(video,acc,acc.fw1:acc.rv4, factor_key=TRUE) %>%
  mutate(video = str_sub(video,-3,-1))
data.forehead <- data %>%
  select(id,forehead.fw1:forehead.rv4) %>%
  gather(video,forehead,forehead.fw1:forehead.rv4) %>%
  mutate(video = str_sub(video,-3,-1))
data.eyes <- data %>%
  select(id,eyes.fw1:eyes.rv4) %>%
  gather(video,eyes,eyes.fw1:eyes.rv4) %>%
  mutate(video = str_sub(video,-3,-1))
data.mouth <- data %>%
  select(id,mouth.fw1:mouth.rv4) %>%
  gather(video,mouth,mouth.fw1:mouth.rv4) %>%
  mutate(video = str_sub(video,-3,-1))
data.chin <- data %>%
  select(id,chin.fw1:chin.rv4) %>%
  gather(video,chin,chin.fw1:chin.rv4) %>%
  mutate(video = str_sub(video,-3,-1))
data.upperchest <- data %>%
  select(id,upperchest.fw1:upperchest.rv4) %>%
  gather(video,upperchest,upperchest.fw1:upperchest.rv4) %>%
  mutate(video = str_sub(video,-3,-1))
data.midchest <- data %>%
  select(id,midchest.fw1:midchest.rv4) %>%
  gather(video,midchest,midchest.fw1:midchest.rv4) %>%
  mutate(video = str_sub(video,-3,-1))
data.lowerchest <- data %>%
  select(id,lowerchest.fw1:lowerchest.rv4) %>%
  gather(video,lowerchest,lowerchest.fw1:lowerchest.rv4) %>%
  mutate(video = str_sub(video,-3,-1))
data.belly <- data %>%
  select(id,belly.fw1:belly.rv4) %>%
  gather(video,belly,belly.fw1:belly.rv4) %>%
  mutate(video = str_sub(video,-3,-1))
data.left <- data %>%
  select(id,left.fw1:left.rv4) %>%
  gather(video,left,left.fw1:left.rv4) %>%
  mutate(video = str_sub(video,-3,-1))
data.right <- data %>%
  select(id,right.fw1:right.rv4) %>%
  gather(video,right,right.fw1:right.rv4) %>%
  mutate(video = str_sub(video,-3,-1))
data.header <- data[1:11]

# Join them all back together. inner_join is smart, it'll join by id AND video.
newdata <- data.header %>%
  inner_join(data.acc,by="id") %>%
  inner_join(data.forehead) %>%
  inner_join(data.eyes) %>%
  inner_join(data.mouth) %>%
  inner_join(data.chin) %>%
  inner_join(data.upperchest) %>%
  inner_join(data.midchest) %>%
  inner_join(data.lowerchest) %>%
  inner_join(data.belly) %>%
  inner_join(data.left) %>%
  inner_join(data.right)

# Let's change the names
oldata <- data
data <- newdata

# Add story and direction variables and split data into videogroups
data$story <- NA
data$direction <- NA
group1 <- filter(data,videogroup=="Group 1")
group2 <- filter(data,videogroup=="Group 2")

# Now define levels for story and direction based on that table above
group1 <- mutate(group1,story = ifelse(video == "fw1","Goldilocks",
                                       ifelse(video == "rv2","Cinderella",
                                            ifelse(video=="fw3","KingMidas","RedRidingHood"))))
group1 <- mutate(group1,direction = ifelse(video == "fw1","forward",
                                           ifelse(video == "rv2","reversed",
                                                  ifelse(video == "fw3","forward","reversed"))))
group2 <- mutate(group2,story = ifelse(video == "fw1","RedRidingHood",
                                       ifelse(video == "rv2","KingMidas",
                                              ifelse(video =="fw3","Cinderella","Goldilocks"))))
group2 <- mutate(group2,direction = ifelse(video == "fw1","forward",
                                       ifelse(video == "rv2","reversed",
                                              ifelse(video == "fw3","forward","reversed"))))



# Join groups back together and view
data <- rbind(group1,group2)
data <- arrange(data,id,video)

# Convert some columns to factors
data <- data %>%
  mutate(hearing = as.factor(hearing)) %>%
  mutate(videogroup = as.factor(videogroup)) %>%
  mutate(aoagroup = as.factor(aoagroup)) %>%
  mutate(languagegroup = as.factor(languagegroup)) %>%
  mutate(maingroup = as.factor(maingroup)) %>%
  mutate(video = as.factor(video)) %>%
  mutate(story = as.factor(story)) %>%
  mutate(direction = as.factor(direction)) %>%
  select(id,participant,hearing,videogroup,aoagroup,languagegroup,maingroup,video,story,
         direction,age,selfrate,signyrs,aoasl,acc,forehead,eyes,mouth,chin,upperchest,
         midchest,lowerchest,belly,left,right)

```
Here's the final, "cleaned-up" dataset that we're going to use for all further analysis. Just the first 10 rows. Notice that there are now story and direction columns, and AOI-only columns, and each participant has 4 observation rows (one observation per story). So, more rows, less columns.  
```{r}
#kable(head(data, n=10)) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
select(data,-participant)
```

# Participant Demographics
Now we can easily get group means.
```{r}
groupmeans <- data %>%
  group_by(maingroup) %>%
  summarize(n = n()/4,
            age = mean(age),
            selfrate = mean(selfrate),
            signyrs = mean(signyrs),
            aoasl = mean(aoasl))
kable(groupmeans, digits=1) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
#groupmeans
```

# Save the dataset for further analysis
We'll save this nice clean dataset now and call it `cleandata.csv`. 
```{r}
write.csv(data,"cleandata.csv",row.names=FALSE)
```

Happy? You can proceed to [Lexical Recall Analysis](02lexicalrecall.nb.html) or [Eye Gaze Analysis](03eyegaze.nb.html).