---
title: "Viewing Space (study1adults)"
author: "Adam Stone, PhD" 
date: '`r format(Sys.Date(), "%m-%d-%Y")`'
output:
  github_document:
    toc: yes
    toc_depth: 2
  html_notebook:
    code_folding: hide
    theme: paper
    highlight: tango
    toc: yes
    toc_depth: 2
    toc_float: yes
    df_print: paged
---

# Get Data
We'll load the raw data from all 8 .xls files. Each file shows the participant's name, group, language, media, and x/y eye gaze coordinates. Here's a quick glimpse of the data and how it's structured after I've processed and cleaned it up. 
```{r, message=FALSE}
# Import packages we'll need.
library(tidyverse)
library(stringr)

# Gather up the files
files <- list.files(pattern = "\\.csv",path="rawdata")
files <- str_c("rawdata/",files)
rawdata <- do.call("rbind", lapply(files, read_csv))

# Clean Up
rawdataoriginal <- rawdata
rawdata <- rawdataoriginal
rawdata <- rawdata %>%
  rename(participant = ParticipantName,
         group = "[Group]Value",
         language = "[Language]Value",
         media = MediaName,
         x = "GazePointX (MCSpx)",
         y = "GazePointY (MCSpx)") %>%
  select(participant,group,language,media,x,y) %>%
  add_column(story=NA,direction=NA) %>%
  mutate(story = case_when(
    str_detect(media,"bears") ~ "bears",
    str_detect(media,"cinderella") ~ "cinderella",   
    str_detect(media,"midas") ~ "midas",
    str_detect(media,"redridinghood") ~ "redridinghood")) %>%
  mutate(direction = case_when(
    str_detect(media,"FW") ~ "forward",
    str_detect(media,"ER") ~ "reversed")) %>%
  mutate(media = str_c(story,direction, sep="_"))
glimpse(rawdata)
```

Now that it's in the right format...it's ridiculously easy to get what we need. For example, here's the mean x and y coordinate for each story for each participant. 

```{r}
means <- rawdata %>%
  group_by(participant,media) %>%
  dplyr::summarize(x = mean(x,na.rm=TRUE),
            y = mean(y,na.rm=TRUE))
means
```

And I can get x or y plots of one participant across 4 stories. Let's do Adam (me?). 

```{r}
adam <- filter(rawdata,participant=="Adam")
ggplot(adam,aes(x=x,y=y,color=media)) + geom_point() + geom_line() + facet_wrap("media",ncol=2,nrow=2) + guides(color="none")
```

