---
title: "Gist (study1adults)"
author: "Adam Stone, PhD" 
date: '`r format(Sys.Date(), "%m-%d-%Y")`'
output:
  github_document:
    toc: yes
    toc_depth: 2
  html_notebook:
    code_folding: hide
    theme: paper
    highlight: tango
    toc: yes
    toc_depth: 2
    toc_float: yes
    df_print: paged
---

# Introduction 
Okay let's look at gist!

# Participants
```{r warning=FALSE}
# Load libraries
library(tidyverse)
library(stringr)
library(lme4)
library(lmerTest)
library(prettydoc)
library(broom)
library(knitr)
library(xtable)
library(kableExtra)
library(viridis)
library(cowplot)
options(knitr.table.format = "html") 

# Import data!
data <- read_csv('cleanpercentdata.csv',col_types = 
                   cols(
                     id = col_integer(),
                     participant = col_character(),
                     hearing = col_character(),
                     videogroup = col_character(),
                     aoagroup = col_character(),
                     languagegroup = col_character(),
                     maingroup = col_character(),
                     video = col_character(),
                     story = col_character(),
                     direction = col_character(),
                     age = col_double(),
                     selfrate = col_double(),
                     signyrs = col_double(),
                     aoasl = col_integer(),
                     acc = col_double(),
                     aoi = col_character(),
                     percent = col_double()
                   ))

# And factorize
data <- data %>%
  mutate(participant = as.factor(participant)) %>%
  mutate(id = as.factor(id)) %>%
  mutate(hearing = as.factor(hearing)) %>%
  mutate(videogroup = as.factor(videogroup)) %>%
  mutate(aoagroup = as.factor(aoagroup)) %>%
  mutate(languagegroup = as.factor(languagegroup)) %>%
  mutate(maingroup = as.factor(maingroup)) %>%
  mutate(video = as.factor(video)) %>%
  mutate(story = as.factor(story)) %>%
  mutate(direction = as.factor(direction)) %>%
  mutate(aoi = as.factor(aoi))

# Remove ASL from the end of MainGroup names
data <- data %>%
  mutate(maingroup = case_when(
    str_detect(maingroup,"DeafNative") ~ "DeafNative",
    str_detect(maingroup,"DeafEarlyASL") ~ "DeafEarly",
    str_detect(maingroup,"DeafLateASL") ~ "DeafLate",
    str_detect(maingroup,"HearingLateASL") ~ "HearingLate",
    str_detect(maingroup,"HearingNoviceASL") ~ "HearingNovice"
  )) %>%
  mutate(maingroup = as.factor(maingroup))

# Set reference levels for maingroup
data$maingroup <- relevel(data$maingroup, ref="DeafNative")

dataoriginal <- data # Save item-level data just in case

# Take out HearingNoviceASL
# data <- data %>%
#   filter(maingroup!="HearingNoviceASL")

# Load awesome function to make correlation tables with stars for significance
# From: https://myowelt.blogspot.co.uk/2008/04/beautiful-correlation-tables-in-r.html
corstarsl <- function(x){ 
require(Hmisc) 
x <- as.matrix(x) 
R <- Hmisc::rcorr(x)$r 
p <- Hmisc::rcorr(x)$P 
## define notions for significance levels; spacing is important.
mystars <- ifelse(p < .001, "***", ifelse(p < .01, "** ", ifelse(p < .05, "* ", " ")))
## trunctuate the matrix that holds the correlations to two decimal
R <- format(round(cbind(rep(-1.11, ncol(x)), R), 2))[,-1] 
## build a new matrix that includes the correlations with their apropriate stars 
Rnew <- matrix(paste(R, mystars, sep=""), ncol=ncol(x)) 
diag(Rnew) <- paste(diag(R), " ", sep="") 
rownames(Rnew) <- colnames(x) 
colnames(Rnew) <- paste(colnames(x), "", sep="") 
## remove upper triangle
Rnew <- as.matrix(Rnew)
Rnew[upper.tri(Rnew, diag = TRUE)] <- ""
Rnew <- as.data.frame(Rnew) 
## remove last column and return the matrix (which is now a data frame)
Rnew <- cbind(Rnew[1:length(Rnew)-1])
return(Rnew) 
}

prob_calc <- function(x){
  odds <- exp(x)
  prob <- odds / (1 + odds)
  return(prob)
}

# # Now collapse eye gaze data to subject-level 
# data <- data %>%
#   group_by(participant,direction,aoi) %>%
#   dplyr::summarize(percent = mean(percent,na.rm=TRUE))
# data[data=="NaN"] <- NA
# 
# # Join subject info with data that's now subject-level
# data <- left_join(data,data.subjectinfo, by=c("participant","direction"))


# But now we need to go back and add in a complete lexical recall dataset, even including those trials that got thrown out in 03eyegaze.nb.html. Because the lexical accuracy data is still good. So let's work on that. 
cleanlexdata <- read_csv('cleandata.csv',col_types = 
                   cols(
                     id = col_integer(),
                     participant = col_character(),
                     hearing = col_character(),
                     videogroup = col_character(),
                     aoagroup = col_character(),
                     languagegroup = col_character(),
                     maingroup = col_character(),
                     video = col_character(),
                     story = col_character(),
                     direction = col_character(),
                     age = col_double(),
                     selfrate = col_double(),
                     signyrs = col_double(),
                     aoasl = col_integer(),
                     acc = col_double(),
                     forehead = col_double(),
                     eyes = col_double(),
                     mouth = col_double(),
                     chin = col_double(),
                     upperchest = col_double(),
                     midchest = col_double(),
                     lowerchest = col_double(),
                     belly = col_double(),
                     left = col_double(),
                     right = col_double(),
                     total = col_double()
                   )) %>%
  mutate(maingroup = case_when(
    str_detect(maingroup,"DeafNative") ~ "DeafNative",
    str_detect(maingroup,"DeafEarlyASL") ~ "DeafEarly",
    str_detect(maingroup,"DeafLateASL") ~ "DeafLate",
    str_detect(maingroup,"HearingLateASL") ~ "HearingLate",
    str_detect(maingroup,"HearingNoviceASL") ~ "HearingNovice"
  )) %>%
  mutate(maingroup = as.factor(maingroup))

# Pull out subject info for later in summary tables
subjectinfo <- data %>%
  dplyr::select(-aoi,-percent,-video,-story,-direction,-acc) %>%
  distinct()

# Participant Characteristics Table (using cleanlexdata because it's more complete)
groupmeans <- cleanlexdata %>%
  ungroup() %>%
  dplyr::select(id,participant,maingroup,age,selfrate,signyrs,aoasl) %>%
  distinct() %>%
  group_by(maingroup) %>%
  dplyr::summarize(n = n(),
            age.m = mean(age),
            age.sd = sd(age),
            selfrate.m = mean(selfrate),
            selfrate.sd = sd(selfrate),
            signyrs.m = mean(signyrs),
            signyrs.sd = sd(signyrs),
            aoasl.m = mean(aoasl),
            aoasl.sd = sd(aoasl)) %>%
  mutate(maingroup =  factor(maingroup, levels = c("DeafNative","DeafEarly","DeafLate",
                                                   "HearingLate","HearingNovice"))) %>%
  arrange(maingroup)    
kable(groupmeans, digits=1) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

Next we fold in the gist data to both datasets. Because the gist data is direction-level we have to summarize both eyegaze and lexrecall datasets. After doing that we can get a quick snapshot of what comprehension looked like across groups. 

```{r}
# Import gist data
gist <- read_csv('gist_indiv.csv', col_types = cols(
  participant = col_character(),
  gist.fw1 = col_integer(),
  gist.rv2 = col_integer(),
  gist.fw3 = col_integer(),
  gist.rv4 = col_integer()
)) %>%
  gather(video, gist, gist.fw1:gist.rv4) %>%
  mutate(video = str_sub(video,6,8))

# Join eye and lex data with gist
data <- data %>%
  left_join(gist, by = c("participant", "video")) %>%
  mutate(maingroup = factor(maingroup, levels = c("DeafNative","DeafEarly","DeafLate",
                                                  "HearingLate","HearingNovice"))) %>%
  mutate(gist = factor(gist, labels = c("No","Yes")))

cleanlexdata <- cleanlexdata %>%
  dplyr::select(id:acc) %>%
  left_join(gist, by = c("participant", "video")) %>%
  mutate(maingroup = factor(maingroup, levels = c("DeafNative","DeafEarly","DeafLate",
                                                  "HearingLate","HearingNovice"))) %>%
  mutate(gistfactor = factor(gist, labels = c("No","Yes"))) %>%
  rename(gist_int = gist,
         gist = gistfactor)

# Bar chart of gist across groups
cleanlexdata %>%
  group_by(maingroup, direction) %>%
  dplyr::summarize(gist = mean(gist_int, na.rm=TRUE)) %>%
  ggplot(aes(x = maingroup, y = gist, fill = direction)) + geom_col(position = "dodge")

cleanlexdata %>%
  dplyr::select(maingroup,direction,gist) %>%
  group_by(maingroup,direction) %>%
  count(gist) %>%
  ungroup() %>%
  spread(gist,n) %>%
  mutate(percent = if_else(!is.na(No), Yes/(No+Yes), 1)) %>%
  dplyr::select(maingroup,direction,percent) %>%
  spread(direction,percent) %>%
  kable(digits=2) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

So we have this weird pattern of DeafLate and HearingLate doing best on the forward stories - even better than DeafNative or DeafEarly! We have a decent pattern for the reversed stories, though. 

# Lexical Recall
Immediately, one would think lexical recall and gist should be correlated, right? Let's check it out. 
```{r}
cleanlexdata %>%
  ggplot(aes(x = gist_int, y = acc, color = direction)) + geom_jitter(width = .1, size = .5) + 
  geom_smooth(method = "lm", size = .75) + facet_grid(direction ~ maingroup) + 
  scale_x_continuous(breaks = c(0, 1)) +
  scale_y_continuous(limits = c(0,1))
```

Hmm. First, let's see if gist changes accuracy, via a LMM with predictors gist, maingroup, direction, and grouping variables participant & story. I think the output is cool...it tells us there are main effects of group (HearingLate and HearingNovice) on accuracy, and there's an interaction between those two groups and gist, meaning if they understand the story, accuracy would go up for those two groups. 

Wha I find weird is reversal is not important here, I wonder if gist and direction are highly correlated and therefore collinear. I ran models collapsing maingroup together, and in those, direction is a significant effect on acc along with gist, so it's probably just loss of power. 

```{r}
gistonacc_lmm <- lmer(acc ~ gist * direction * maingroup + (1|id) + (1|story), data = cleanlexdata)
summary(gistonacc_lmm)
```

Is there an effect of maingroup and/or direction on gist? Let's do a LMM but with a logit-link function since it's a binary outcome. 

Nothing significant here, mainly because  most people got gist correctly for forward stories so it's not a very normal thingy... or this model is just screwy. The reversed-only model below gets it better. 

```{r}
gist_glmm <- glmer(gist ~ direction * maingroup + (1|id) + (1|story), data = cleanlexdata, family=binomial (link="logit"))
summary(gist_glmm)
```

# AoA effects for Reversed Stories on Gist!
The coefficients are log odds. HearingNovice and Direction are significant main effects. I converted them to probability by hand (odds = exp(coef), then probability = odds / 1+odds). From [Sebastian's blog](https://sebastiansauer.github.io/convert_logit2prob/). [And a good SO post here.](https://stackoverflow.com/questions/41384075/r-calculate-and-interpret-odds-ratio-in-logistic-regression) The intercept is around 1, so the probability of getting it right as a DeafNative is 100%. (If I'm interpreting this right). Adding up the coefficients:

What about reversed stories only? Doing it again here.

1. DeafNative probability of getting gist correctly (the intercept): 79%
1. DeafEarly probability: 33.5% (significant diff from DeafNative, p = 0.039)
1. DeafLate probability: 40% (not significantly diff from DeafNative, p = 0.079)
1. HearingLate probability: 26% (significantly diff from DeafNative, p = 0.014)
1. HearingNovice probability: 14% (significantly diff from DeafNative, p = 0.004)

```{r}
gist_glmm_r <- glmer(gist ~ maingroup + (1|id) + (1|story), data = filter(cleanlexdata, direction=="reversed"), family=binomial (link="logit"))
summary(gist_glmm_r)
```

I ran posthoc contrasts here. Tells us HearingNovice is driving the differences  (HearingNovice significantly different from DeafNative at 0.0362). 

```{r}
library(multcomp)
summary(glht(gist_glmm_r, linfct = mcp(maingroup = "Tukey")))
```

# Heat Maps
I'm going to focus on reversed stories only for now. Can we see differences in eye gaze between those who "got" the reversed story and those who didn't?

```{r}
eye_heat <- data %>%
  filter(direction == "reversed") %>%
  filter(aoi != "left" & aoi != "right" & aoi != "facechest" & aoi != "face" & aoi != "chest") %>%
  group_by(gist, maingroup, participant, aoi) %>%
  dplyr::summarize(percent = mean(percent, na.rm=TRUE)) %>%
  group_by(gist, maingroup, aoi) %>%
  dplyr::summarize(percent = mean(percent, na.rm=TRUE)) %>%
#  spread(aoi,percent) %>%
#  filter(!is.na(aoi)) %>%
  mutate(aoi = factor(aoi,levels=c("belly","lowerchest","midchest",
                                   "upperchest","chin","mouth","eyes","forehead")))

eye_heat %>%
  group_by(gist,aoi) %>%
  dplyr::summarize(percent = mean(percent, na.rm=TRUE)) %>%
  ggplot(aes(x = gist, y = aoi)) +
  geom_tile(aes(fill=percent),color="lightgray",na.rm=TRUE) + 
  scale_fill_viridis(option = "viridis", direction=-1, limits = c(0,1)) +
  theme(axis.text.x=element_text(angle=45,hjust=1)) + 
  ylab("") + xlab("") + ggtitle("Heat Map for Reversed Stories Only")

eye_heat %>%
  ggplot(aes(x = gist, y = aoi)) +
  geom_tile(aes(fill=percent),color="lightgray",na.rm=TRUE) + 
  scale_fill_viridis(option = "viridis", direction=-1, limits = c(0,1)) +
  theme(axis.text.x=element_text(angle=45,hjust=1)) + facet_grid(.~maingroup) +
  ylab("") + xlab("") + ggtitle("Heat Map for Reversed Stories Only")

eye_heat %>%
  ggplot(aes(x = maingroup, y = aoi)) +
  geom_tile(aes(fill=percent),color="lightgray",na.rm=TRUE) + 
  scale_fill_viridis(option = "viridis", direction=-1, limits = c(0,1)) +
  theme(axis.text.x=element_text(angle=45,hjust=1)) + facet_grid(.~gist) +
  ylab("") + xlab("") + ggtitle("Heat Map for Reversed Stories Only")
```

# Gist & Gaze Modeling

That's cool, right? Anyway so I am now going to try to make models. I was thinking...should eye gaze behavior predict gist, or gist predict eye behavior? Which direction could this work? Gist is measured after eye gaze, so it sounds like eye gaze should predict gist. 

But that can't be right. Eye gaze is a function of whether one can understand or comprehend the story. So gist should predict eye gaze changes. *Aside: I've read criticism of the use "predict" in psychology and that it's better characterized as "associated with"...we can think about that in the write-up.* 

So a basic model would be (A) gist -> eye gaze. We could predict group differences such that (B) gist X group -> eye gaze. Let's give it a shot. Breaking it down by AOI, and checking both A and B. **Again, reversed stories only. We can add forward stories back in later, maybe, or analyze separately**

## Eye AOI
(A) No effect of gist. 
(B) No effect of gist or maingroup.
```{r}
data_r <- data %>%
  filter(direction=="reversed") %>%
  spread(aoi,percent)

eye_lm <- lmer(eyes ~ gist + (1|id) + (1|story), data = data_r)
summary(eye_lm)

eye_lm_mg <- lmer(eyes ~ gist * maingroup + (1|id) + (1|story), data = data_r)
summary(eye_lm_mg)
```

## Mouth AOI
(A) No effect of gist. 
(B) No effect of gist or maingroup. 
```{r}
mouth_lm <- lmer(mouth ~ gist + (1|id) + (1|story), data = data_r)
summary(mouth_lm)

mouth_lm_mg <- lmer(mouth ~ gist * maingroup + (1|id) + (1|story), data = data_r)
summary(mouth_lm_mg)
```

## Chin AOI
(A) No effect of gist. 
(B) No effect of gist or maingroup. 
```{r}
chin_lm <- lmer(chin ~ gist + (1|id) + (1|story), data = data_r)
summary(chin_lm)

chin_lm_mg <- lmer(chin ~ gist * maingroup + (1|id) + (1|story), data = data_r)
summary(chin_lm_mg)
```

## FaceChest Ratio
(A) No effect of gist. 
(B) No effect of gist. Main effect of HearingNovice (much lower facechest ratio) 
```{r}
fcr_lm <- lmer(facechest ~ gist + (1|id) + (1|story), data = data_r)
summary(fcr_lm)

fcr_lm_mg <- lmer(facechest ~ gist * maingroup + (1|id) + (1|story), data = data_r)
summary(fcr_lm_mg)
```

## MouthEye Ratio (+1 = Mouth, -1 = Eye)
(A) No effect of gist. 
(B) No effect of gist. Main effect of HearingLate (much lower moutheye ratio (meaning more looking at eye)). 
```{r}
me_lm <- lmer(moutheye ~ gist + (1|id) + (1|story), data = data_r)
summary(me_lm)

me_lm_mg <- lmer(moutheye ~ gist * maingroup + (1|id) + (1|story), data = data_r)
summary(me_lm_mg)
```

I want to make and see DeafLate's mouth AOI model...and eyes and chin. Nothing.  
```{r}
deaflate <- data %>% filter(maingroup == "DeafLate") %>% spread(aoi,percent)

deaflate_m <- lmer(chin ~ gist * direction + (1|id) + (1|story), data = deaflate)
summary(deaflate_m)
```

# Correlations
Now I wanna do correlations. We've got eye gaze and lex recall and gist in the `data` object now, right? Okay I'll break it down to lexrecall, gist, and have forward/reversed flavors for each. 

GIST and FORWARD (is this fair to even compute? not normal...85 out of 98 forward stories were correct gist.)
```{r  rows.print=20}
gaze_gist_fw <- data %>% 
  filter(direction == "forward") %>%
  spread(aoi,percent) %>%
  dplyr::select(-(id:acc)) %>%
  mutate(gist = case_when(
    gist == "Yes" ~ 1,
    gist == "No" ~ 0
  ))

corstarsl(gaze_gist_fw)
```

GIST and REVERSED. No correlations. 
```{r rows.print=20}
gaze_gist_rv <- data %>% 
  filter(direction == "reversed") %>%
  spread(aoi,percent) %>%
  dplyr::select(-(id:acc)) %>%
  mutate(gist = case_when(
    gist == "Yes" ~ 1,
    gist == "No" ~ 0
  ))

corstarsl(gaze_gist_rv)
```

LEX RECALL and FORWARD. Only one correlation - forehead is negatively correlated with accuracy. There was some increased forehead looking in reversed stories that we didn't see in forward stories. But forward stories has really tiny forehead percentages. May not be good to report it.
```{r rows.print=20}
gaze_lex_fw <- data %>% 
  filter(direction == "forward") %>%
  spread(aoi,percent) %>%
  dplyr::select(-(id:aoasl)) %>%
  dplyr::select(-gist)

corstarsl(gaze_lex_fw)
```

LEX RECALL and REVERSED. Face-looking correlated with increased accuracy in reversed stories. Nice. r = 0.23, p <= 0.05. 
```{r rows.print=20}
gaze_lex_rv<- data %>% 
  filter(direction == "reversed") %>%
  spread(aoi,percent) %>%
  dplyr::select(-(id:aoasl)) %>%
  dplyr::select(-gist)

corstarsl(gaze_lex_rv)
```

# Moutheye Ratio
Checking out moutheye ratio. Okay I think this is interesting. HearingLate has much more variance for moutheye, quite low (a lot of eye looking). 
```{r}
me_data <- data %>%
  dplyr::select(-aoi, -percent, -gist) %>%
  distinct()

ggplot(me_data, aes(x = maingroup, y = moutheye, fill = direction)) + geom_boxplot()
```

Is there an AoA effect on moutheye ratio? Nope...
```{r}
me_aoa <- lmer(moutheye ~ maingroup * direction + (1|id) + (1|story), data = me_data)
summary(me_aoa)
```

AoA effect for reversed stories only? Yes, HearingLate. 

```{r}
me_aoa_r <- lmer(moutheye ~ maingroup + (1|id) + (1|story), data = filter(me_data, direction=="reversed"))
summary(me_aoa_r)
```
