---
title: "The Last Data Analysis to Rule Them All? (study1adults)"
author: "Adam Stone, PhD"
date: '`r format(Sys.Date(), "%m-%d-%Y")`'
output:
  html_notebook:
    code_folding: hide
    highlight: tango
    theme: paper
    toc: yes
    toc_depth: 2
    toc_float: yes
---

# Putting It All Back Together

Throughout all the data analysis we've done, the datasets have become more fragmented - lexical recall, gist, and eye tracking datasets. I want to put them all together in one whole dataset again so we can perform some analyses more efficiently (particularly correlations). The only thing I need to remember is we'll have a new column called `eye_exclude` and if it is set to `TRUE` it means we can't include that row in any analysis relating to eye gaze (usually because that trial was less than 25% looking). 

```{r smash data together, message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(lme4)
library(lmerTest)
library(scales)
library(viridis)
library(agricolae) 
library(GGally)

# Load lex and eye data
cleanlexdata <- read_csv("cleandata.csv") %>%
  select(-(forehead:total))

cleaneyedata <- read_csv("cleanpercentdata.csv") %>%
  spread(aoi,percent) %>%
  add_column(eye_exclude = FALSE)

# What rows were removed from the eye data back in 03eyegaze? Let's add back in
# With a new column - eye_exclude
removed <- anti_join(cleanlexdata, cleaneyedata) %>%
  add_column(eye_exclude = TRUE)
eyelexdata <- bind_rows(cleaneyedata, removed)

# Load gist data
gist <- read_csv('gist_indiv.csv', col_types = cols(
  participant = col_character(),
  gist.fw1 = col_integer(),
  gist.rv2 = col_integer(),
  gist.fw3 = col_integer(),
  gist.rv4 = col_integer()
)) %>%
  gather(video, gist, gist.fw1:gist.rv4) %>%
  mutate(video = str_sub(video,6,8))

# Presto, our full reunified dataset - 'fulldata'
# But I want to remove columns I don't want anymore and will recalculate later
fulldata <- left_join(eyelexdata, gist) %>%
  select(-moutheye, -facechest, -face, -chest)
```

# Group Changes and Participant Tables

We have some changes to make to the groups. First, fix Josh as learning ASL when he was 6. Next, drop the DeafNative Group and reclassify all who learned ASL < 3.9 as DeafEarly and ASL => 4.0 as DeafLate. 

```{r groups and participant tables}
# Change Josh's AoASL to 6
fulldata <- fulldata %>%
  mutate(aoasl = as.double(aoasl)) %>%
  mutate(aoasl = case_when(
    participant == "Josh" ~ 6,
    TRUE ~ aoasl
  ))

# Reclassify Groups
fulldata <- fulldata %>%
  mutate(maingroup = case_when(
    hearing == "Deaf" & aoasl < 4 ~ "DeafEarly",
    hearing == "Deaf" & aoasl >= 4 ~ "DeafLate",
    maingroup == "HearingLateASL" ~ "HearingLate",
    maingroup == "HearingNoviceASL" ~ "HearingNovice"
  ))

# Create Participant Demographics Table
participant_info <- fulldata %>%
  select(-(acc:gist)) %>%
  select(-(video:direction)) %>%
  distinct() %>% 
  group_by(maingroup) %>%
  summarise(n = n(),
            age_mean = mean(age),
            age_sd = sd(age),
            aoasl_mean = mean(aoasl),
            aoasl_sd = sd(aoasl),
            signyrs_mean = mean(signyrs),
            signyrs_sd = sd(signyrs),
            selfrate_mean = mean(selfrate),
            selfrate_sd = sd(selfrate)) %>%
  ungroup() %>%
  mutate_if(is.double, funs(round(., 2))) %>%
  mutate(age = paste(age_mean, "±", age_sd, sep = " "),
         aoasl = paste(aoasl_mean, "±", aoasl_sd, sep = " "),
         signyrs = paste(signyrs_mean, "±", signyrs_sd, sep = " "),
         selfrate = paste(selfrate_mean, "±", selfrate_sd, sep = " ")) %>%
  select(-(age_mean:selfrate_sd))

participant_info
```
## Participant ANOVAs 

Below are the ANOVA outputs for participant demographics, and LSDs for each. 

Participants' age
```{r subject anova age, echo=FALSE, warning=FALSE}
# Make the participant data frame
p_anovas <- fulldata %>%
  select(-(acc:gist)) %>%
  select(-(video:direction)) %>%
  distinct() %>%
  select(maingroup:aoasl)

# Age ANOVA
p_age <- aov(age ~ maingroup, data = p_anovas)
summary(p_age)
p_age_lsd <- LSD.test(p_age, "maingroup", group = FALSE)$comparison
p_age_lsd
```


Participants' AoASL
```{r subject anova aoasl, echo=FALSE, warning=FALSE}
# AoASL ANOVA
p_aoasl <- aov(aoasl ~ maingroup, data = p_anovas)
summary(p_aoasl)
p_aoasl_lsd <- LSD.test(p_aoasl, "maingroup", group = FALSE)$comparison
p_aoasl_lsd
```

Participants' Sign Yrs
```{r subject anova signyrs, echo=FALSE, warning=FALSE}
# Signyrs ANOVA
p_signyrs <- aov(signyrs ~ maingroup, data = p_anovas)
summary(p_signyrs)
p_signyrs_lsd <- LSD.test(p_signyrs, "maingroup", group = FALSE)$comparison
p_signyrs_lsd
```

Participants' Self-Rating
```{r subject anova selfrate, echo=FALSE, warning=FALSE}
# Selfrate ANOVA
p_selfrate <- aov(selfrate ~ maingroup, data = p_anovas)
summary(p_selfrate)
p_selfrate_lsd <- LSD.test(p_selfrate, "maingroup", group = FALSE)$comparison
p_selfrate_lsd
```

# Gist & Lexical Recall Data

## Tables & Charts

Let's generate a table for lexical recall and gist for forward vs. reversed stories. 

```{r lex & gist table}
lexgist_info <- fulldata %>%
  group_by(maingroup, direction) %>%
  summarise(lex_mean = mean(acc, na.rm = TRUE),
            lex_sd = sd(acc, na.rm = TRUE),
            gist_mean = mean(gist),
            gist_sd = sd(gist)) %>%
  ungroup() %>%
  mutate_if(is.double, funs(round(., 2))) %>% 
  mutate(lex = paste(lex_mean, "±", lex_sd, sep = " "),
         gist = paste(gist_mean, "±", gist_sd, sep = " ")) %>%
  select(-(lex_mean:gist_sd)) %>%
  gather(metric, value, lex:gist) %>%
  unite("metric", c(metric, direction), sep = "_") %>%
  spread(metric, value) %>%
  print()
```

And then bar charts too after that with error bars. 

```{r lex and gist bar charts}
# Gist bar chart
gist_bar <- fulldata %>% select(participant, maingroup, direction, gist) %>%
  group_by(maingroup, participant, direction) %>%
  summarise(gist = mean(gist)) %>%
  group_by(maingroup, direction) %>%
  summarise(mean = mean(gist),
            sd = sd(gist),
            count = n(),
            se = sd/sqrt(count))

ggplot(gist_bar, aes(x = maingroup, y = mean, fill = direction)) +
  geom_bar(stat = "identity", position = position_dodge()) + 
  geom_errorbar(aes(ymin = mean-se, ymax = mean+se), position = position_dodge(0.9), width = 0.5) +
  labs(title = "Story Comprehension (Gist)", subtitle = "Error bars represent SE", x = "", y = "mean gist") +
  scale_y_continuous(labels = percent, limits = c(0,1))

# Lex bar chart
lex_bar <- fulldata %>% select(participant, maingroup, direction, acc) %>%
  group_by(maingroup, participant, direction) %>%
  summarise(acc = mean(acc, na.rm = TRUE)) %>%
  group_by(maingroup, direction) %>%
  summarise(mean = mean(acc, na.rm = TRUE),
            sd = sd(acc, na.rm = TRUE),
            count = n(),
            se = sd/sqrt(count))

ggplot(lex_bar, aes(x = maingroup, y = mean, fill = direction)) +
  geom_bar(stat = "identity", position = position_dodge()) + 
  geom_errorbar(aes(ymin = mean-se, ymax = mean+se), position = position_dodge(0.9), width = 0.5) +
  labs(title = "Lexical Recall", subtitle = "Error bars represent SE", x = "", y = "mean accuracy") +
  scale_y_continuous(labels = percent, limits = c(0,1)) +
  geom_hline(yintercept = .5, linetype = "dotted") +
  coord_cartesian(ylim = c(.5,1))
```

## ANOVA Plan
Next, we're going to do ANOVAs and ANCOVAs. We'll always do it in this order. The first three ANOVAs will be followed by LSD of the four maingroups with *uncorrected p-values.* 

1. ANOVA with factors MainGroup & Direction
2. ANOVA with factor MainGroup, for Forward only
3. ANOVA with factor MainGroup, for Reverse only 
4. ANCOVA with factor Direction, and covariate AoASL and Age
5. Regression with variables AoASL and Age, for Forward only
6. Regression with variables AoASL and Age, for Reverse only. 

I did not include Age as a covariate in the first 3 ANOVAs because they did not add to or change the model in any significant way. 

## Gist ANOVAs

1. ANOVA with factors MainGroup & Direction.

```{r gist anova1, echo=FALSE, results = 'markup'}
# First let's make the participant-level dataset with which we'll do our ANCOVAs. 
participant_data <- fulldata %>%
  group_by(maingroup, participant, direction) %>%
  mutate(gist = mean(gist, na.rm = TRUE),
         acc = mean(acc, na.rm = TRUE)) %>%
  select(id, participant, hearing, maingroup, direction, age, aoasl, acc, gist) %>%
  distinct()

# Gist ANOVA 1
gist_aov1 <- aov(gist ~ maingroup * direction, data = participant_data)
summary(gist_aov1)
gist_lsd1 <- LSD.test(gist_aov1, "maingroup", group = FALSE)
gist_lsd1$comparison
```

2. ANOVA with factor MainGroup, for Forward only. Also a Kruskal-Wallis test. And Chi-Sq too. 

```{r gist anova2, echo=FALSE, results = 'markup'}
# Gist ANOVA 2
gist_aov2 <- aov(gist ~ maingroup, data = filter(participant_data, direction == "forward"))
summary(gist_aov2)
gist_lsd2 <- LSD.test(gist_aov2, "maingroup", group = FALSE)
gist_lsd2$comparison

# KW Non-parametric test (like one-way ANOVA)
kruskal.test(gist ~ maingroup, data = as.matrix(filter(participant_data, direction == "forward")))

# Chi Sq
gist_chisq_fw <- participant_data %>%
  ungroup() %>%
  filter(direction == "forward") %>%
  select(maingroup, gist) %>%
  group_by(maingroup, gist) %>%
  summarise(count = n()) %>%
  spread(gist, count) %>%
  rename(none = "0",
         one = "0.5",
         both = "1")

gist_chisq_fw[is.na(gist_chisq_fw)] <- 0L
gist_chisq_fw <- cbind(gist_chisq_fw[,"none"], gist_chisq_fw[,"one"], gist_chisq_fw[,"both"])
chisq.test(gist_chisq_fw)
```

3. ANOVA with factor MainGroup, for Reverse only. 

```{r gist anova3, echo=FALSE, results = 'markup'}
# Gist ANOVA 3
gist_aov3 <- aov(gist ~ maingroup, data = filter(participant_data, direction == "reversed"))
summary(gist_aov3)
gist_lsd3 <- LSD.test(gist_aov3, "maingroup", group = FALSE)
gist_lsd3$comparison

# KW Non-parametric test (like one-way ANOVA)
kruskal.test(gist ~ maingroup, data = as.matrix(filter(participant_data, direction == "reversed")))

# Chi Sq
gist_chisq_rv <- participant_data %>%
  ungroup() %>%
  filter(direction == "reversed") %>%
  select(maingroup, gist) %>%
  group_by(maingroup, gist) %>%
  summarise(count = n()) %>%
  spread(gist, count) %>%
  rename(none = "0",
         one = "0.5",
         both = "1")

gist_chisq_rv[is.na(gist_chisq_rv)] <- 0L
gist_chisq_rv <- cbind(gist_chisq_rv[,"none"], gist_chisq_rv[,"one"], gist_chisq_rv[,"both"])
chisq.test(gist_chisq_rv)
```

4. ANCOVA with factor Direction, and variables AoASL and Age.

```{r gist anova4, echo=FALSE, results = 'markup'}
# Gist ANOVA 4
gist_aov4 <- aov(gist ~ direction * aoasl * age, data = participant_data)
summary(gist_aov4)
```

5. Regression with variables Age and AoASL, for Forward only

```{r gist anova5, echo=FALSE, results = 'markup'}
# Gist ANOVA 5
gist_aov5 <- lm(gist ~ aoasl * age, data = filter(participant_data, direction == "forward"))
summary(gist_aov5)
```

6. Regression with variables AoASL and Age, for Reverse only. 

```{r gist anova6, echo=FALSE, results = 'markup'}
# Gist ANOVA 6
gist_aov6 <- lm(gist ~ aoasl * age, data = filter(participant_data, direction == "reversed"))
summary(gist_aov6)
```

## Lexical Recall ANOVAs

1. ANOVA with factors MainGroup & Direction.

```{r acc anova1, echo=FALSE, results = 'markup'}
# Lexical Recall ANOVA 1
acc_aov1 <- aov(acc ~ maingroup * direction, data = participant_data)
summary(acc_aov1)
acc_lsd1 <- LSD.test(acc_aov1, "maingroup", group = FALSE)
acc_lsd1$comparison
```

2. ANOVA with factor MainGroup, for Forward only.

```{r acc anova2, echo=FALSE, results = 'markup'}
# Lexical Recall ANOVA 2
acc_aov2 <- aov(acc ~ maingroup, data = filter(participant_data, direction == "forward"))
summary(acc_aov2)
acc_lsd2 <- LSD.test(acc_aov2, "maingroup", group = FALSE)
acc_lsd2$comparison
```

3. ANOVA with factor MainGroup, for Reverse only. 

```{r acc anova3, echo=FALSE, results = 'markup'}
# Lexical Recall ANOVA 3
acc_aov3 <- aov(acc ~ maingroup, data = filter(participant_data, direction == "reversed"))
summary(acc_aov3)
acc_lsd3 <- LSD.test(acc_aov3, "maingroup", group = FALSE)
acc_lsd3$comparison
```

4. ANCOVA with factor Direction, and covariate AoASL and Age.

```{r acc anova4, echo=FALSE, results = 'markup'}
# Lexical Recall ANOVA 4
acc_aov4 <- aov(acc ~ direction * aoasl * age, data = participant_data)
summary(acc_aov4)
```

5. Regression with variables Age and AoASL, for Forward only

```{r acc anova5, echo=FALSE, results = 'markup'}
# Lexical Recall ANOVA 5
acc_aov5 <- lm(acc ~ aoasl * age, data = filter(participant_data, direction == "forward"))
summary(acc_aov5)
```

6. Regression with variables AoASL and Age, for Reverse only. 

```{r acc anova6, echo=FALSE, results = 'markup'}
# Lexical Recall ANOVA 6
acc_aov6 <- lm(acc ~ aoasl * age, data = filter(participant_data, direction == "reversed"))
summary(acc_aov6)
```


# AoA Correlations

Next, we want to look at correlations between AoA and Gist, and betwen AoA and Lexical Recall. Rain asked for forward and reversed separately (1) deaf only, (2) hearing only, and (3) both. Let's make it work. 

```{r repackage data for correlations, message=FALSE, warning=FALSE}
# Let's make participant-level data, and have forward/reversed in separate columns
lexgist_data <- fulldata %>%
  group_by(maingroup, participant, direction) %>%
  mutate(gist = mean(gist, na.rm = TRUE),
         lex = mean(acc, na.rm = TRUE)) %>%
  ungroup() %>%
  select(maingroup, participant, hearing, direction, aoasl, signyrs, age, gist, lex) %>%
  distinct() %>%
  gather(metric, value, gist:lex) %>%
  unite(metricvalue, c(metric, direction), sep = "_") %>%
  spread(metricvalue, value) %>%
  select(-participant, -maingroup)

lexgist_deaf <- lexgist_data %>% filter(hearing == "Deaf") %>% select(-hearing)
lexgist_hearing <- lexgist_data %>% filter(hearing == "Hearing") %>% select(-hearing)
lexgist_all <- lexgist_data %>% select(-hearing)

# Load awesome function to make correlation tables with stars for significance
# From: https://myowelt.blogspot.co.uk/2008/04/beautiful-correlation-tables-in-r.html
corstarsl <- function(x){ 
require(Hmisc) 
x <- as.matrix(x) 
R <- Hmisc::rcorr(x)$r 
p <- Hmisc::rcorr(x)$P 
## define notions for significance levels; spacing is important.
mystars <- ifelse(p < .001, "***", ifelse(p < .01, "** ", ifelse(p < .05, "* ", " ")))
## trunctuate the matrix that holds the correlations to two decimal
R <- format(round(cbind(rep(-1.11, ncol(x)), R), 2))[,-1] 
## build a new matrix that includes the correlations with their apropriate stars 
Rnew <- matrix(paste(R, mystars, sep=""), ncol=ncol(x)) 
diag(Rnew) <- paste(diag(R), " ", sep="") 
rownames(Rnew) <- colnames(x) 
colnames(Rnew) <- paste(colnames(x), "", sep="") 
## remove upper triangle
Rnew <- as.matrix(Rnew)
Rnew[upper.tri(Rnew, diag = TRUE)] <- ""
Rnew <- as.data.frame(Rnew) 
## remove last column and return the matrix (which is now a data frame)
Rnew <- cbind(Rnew[1:length(Rnew)-1])
return(Rnew) 
}

# Correlations for Deaf
print("DEAF Correlations - Pearson's r")
#corstarsl(lexgist_deaf)
Hmisc::rcorr(as.matrix(lexgist_deaf))$r
print("DEAF Correlations - P-values")
Hmisc::rcorr(as.matrix(lexgist_deaf))$P
cat(paste("","\n",""))

# Correlations for Hearing
print("HEARING Correlations - Pearson's r")
#corstarsl(lexgist_hearing)
Hmisc::rcorr(as.matrix(lexgist_hearing))$r
print("HEARING Correlations - P-values")
Hmisc::rcorr(as.matrix(lexgist_hearing))$P
cat(paste("","\n",""))

# Correlations for All
print("ALL Correlations - Pearson's r")
#corstarsl(lexgist_all)
Hmisc::rcorr(as.matrix(lexgist_all))$r
print("ALL Correlations - P-values")
Hmisc::rcorr(as.matrix(lexgist_all))$P

```

I'm also including nicely formatted tables with *** indicators of significance for quick referencing. Order: Deaf, Hearing, All. 

```{r pretty correlations}
corstarsl(lexgist_deaf)
corstarsl(lexgist_hearing)
corstarsl(lexgist_all)
```

## Scatterplot of Correlations
Let's visualize what's happening with the correlations here. 
```{r correlation charts, fig.height=12, fig.width=12, message=FALSE, warning=FALSE}
ggpairs(lexgist_data, columns = c(2:8), aes(color = hearing))
```


# Eye Gaze Data

Now eye gaze data. Boxplots first. Also here, we're renaming "chin" to "neck" because that's what it actually is! But we also have to fix all NA's in the percentages to zeros, becuase that's what they actually are. 
```{r eye gaze boxplot}
# rename chin to neck
fulldata <- fulldata %>%
  rename(neck = chin) %>%
  gather(aoi, percent, belly:upperchest)

# Fix all NA's in Percent column to 0
fixpercent <- fulldata$percent
fulldata$percent <- coalesce(fixpercent, 0)
fulldata <- fulldata %>%
  spread(aoi, percent)

fulldata %>%
  filter(eye_exclude == FALSE) %>%
  select(direction, belly:upperchest) %>%
  gather(aoi, percent, belly:upperchest) %>%
  ggplot(aes(x = aoi, y = percent, fill = direction)) + geom_boxplot()
```

But let's try error charts too! Instead of boxplots.

```{r eye gaze error bar chart}
fulldata_error <- fulldata %>%
  filter(eye_exclude == FALSE) %>%
  gather(aoi, percent, belly:upperchest) %>%
  group_by(id, direction, aoi) %>%
  summarise(percent = mean(percent, na.rm = TRUE)) %>%
  ungroup() %>%
  distinct() %>%
  group_by(direction, aoi) %>%
  summarise(mean = mean(percent, na.rm = TRUE),
            sd = sd(percent, na.rm = TRUE),
            count = n(),
            se = sd/sqrt(count))
fulldata_error$aoi <- fct_relevel(fulldata_error$aoi, c("forehead","eyes","mouth","neck","upperchest",
                                                        "midchest","lowerchest","belly","left","right"))

fulldata_error %>%
  ggplot(aes(x = aoi, y = mean, fill = direction)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = mean-se, ymax = mean+se), position = position_dodge(0.9), width = 0.5) +
  labs(title = "Eye Gaze Behavior", subtitle = "Error bars represent SE", x = "", y = "percent looking") +
  scale_y_continuous(labels = percent, limits = c(0,.75))
```

## Big Three-Way ANOVA
Now we're going to try a three-way Group x Direction x AOI ANOVA with the top 3 AOIs (Eyes, Mouth, Neck)

```{r big anova aoi3, echo=FALSE, results = 'markup'}
aoi3 <- c('eyes','mouth','neck')

# Organize the data by those AOIs (but again we need participant-level first)
aoi3_data <- fulldata %>%
  filter(eye_exclude == FALSE) %>%
  select(id, maingroup, hearing, aoasl, age, direction, belly:upperchest) %>%
  gather(aoi, percent, belly:upperchest) %>%
  group_by(id, direction, aoi) %>%
  mutate(percent = mean(percent, na.rm = TRUE)) %>%
  ungroup() %>%
  distinct() %>%
  filter(aoi %in% aoi3) %>%
  spread(aoi, percent)

# Limit to eyes mouth neck
big_aov_aoi3 <- aoi3_data %>%
  gather(aoi, percent, eyes:neck) %>%
  aov(percent ~ maingroup * direction * aoi, data = .)

summary(big_aov_aoi3)

big_aov_aoi3_lsd <- LSD.test(big_aov_aoi3, "aoi", group = FALSE)
big_aov_aoi3_lsd$comparison
```
## Interactions Visualization
So we have significant maingroup:aoi and direction:aoi interactions. Let's try to visualize what can be driving these. We can go back to the SEM chart but break it down...

*The error bars are not 100% accurate, I took a quick-n-easy way around*

```{r}
aoi3_interactions <- fulldata %>%
  filter(eye_exclude == FALSE) %>%
  select(id, participant, maingroup, direction, eyes, mouth, neck) %>%
  gather(aoi, percent, c(eyes, mouth, neck)) %>%
  group_by(maingroup, direction, aoi) %>%
  summarise(mean = mean(percent, na.rm = TRUE),
            sd = sd(percent, na.rm = TRUE),
            count = n()/2,
            se = sd/sqrt(count))
# I need to first collapse across stories for each participant...here I didn't. Must fix later.

aoi3_interactions %>%
  ggplot(aes(x = aoi, y = mean, fill = direction)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = mean-se, ymax = mean+se), position = position_dodge(0.9), width = 0.5) +
  labs(title = "Eye Gaze Behavior", subtitle = "Error bars represent SE", x = "", y = "percent looking") +
  scale_y_continuous(labels = percent) +
  facet_wrap("maingroup")

aoi3_interactions %>%
  ggplot(aes(x = aoi, y = mean, fill = maingroup)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = mean-se, ymax = mean+se), position = position_dodge(0.9), width = 0.5) +
  labs(title = "Eye Gaze Behavior", subtitle = "Error bars represent SE", x = "", y = "percent looking") +
  scale_y_continuous(labels = percent) +
  facet_wrap("direction")

aoi3_interactions %>%
  ggplot(aes(x = maingroup, y = mean, fill = direction)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = mean-se, ymax = mean+se), position = position_dodge(0.9), width = 0.5) +
  labs(title = "Eye Gaze Behavior", subtitle = "Error bars represent SE", x = "", y = "percent looking") +
  scale_y_continuous(labels = percent) +
  facet_wrap("aoi")

aoi3_interactions %>%
  ggplot(aes(x = direction, y = mean, fill = maingroup)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = mean-se, ymax = mean+se), position = position_dodge(0.9), width = 0.5) +
  labs(title = "Eye Gaze Behavior", subtitle = "Error bars represent SE", x = "", y = "percent looking") +
  scale_y_continuous(labels = percent) +
  facet_wrap("aoi")
```


## All the AOI means
Let's get tables of our means here, in various configurations

```{r aoi raw, echo = FALSE, results = 'markup'}
aoi3_data %>%
  group_by(maingroup, direction) %>%
  summarise(eyes = mean(eyes, na.rm = TRUE),
            mouth = mean(mouth, na.rm = TRUE),
            neck = mean(neck, na.rm = TRUE))

aoi3_data %>%
  group_by(maingroup, direction) %>%
  summarise(eyes = mean(eyes, na.rm = TRUE),
            mouth = mean(mouth, na.rm = TRUE),
            neck = mean(neck, na.rm = TRUE)) %>%
  group_by(maingroup) %>%
  summarise(eyes = mean(eyes, na.rm = TRUE),
            mouth = mean(mouth, na.rm = TRUE),
            neck = mean(neck, na.rm = TRUE))  

aoi3_data %>%
  group_by(maingroup, direction) %>%
  summarise(eyes = mean(eyes, na.rm = TRUE),
            mouth = mean(mouth, na.rm = TRUE),
            neck = mean(neck, na.rm = TRUE)) %>%
  group_by(direction) %>%
  summarise(eyes = mean(eyes, na.rm = TRUE),
            mouth = mean(mouth, na.rm = TRUE),
            neck = mean(neck, na.rm = TRUE))  

aoi3_data %>%
  group_by(maingroup, direction) %>%
  summarise(eyes = mean(eyes, na.rm = TRUE),
            mouth = mean(mouth, na.rm = TRUE),
            neck = mean(neck, na.rm = TRUE)) %>%
  group_by(maingroup) %>%
  summarise(eyes = mean(eyes, na.rm = TRUE),
            mouth = mean(mouth, na.rm = TRUE),
            neck = mean(neck, na.rm = TRUE)) %>%
  ungroup() %>% 
  gather(aoi, percent, eyes:neck) %>%
  group_by(aoi) %>%
  summarise(percent = mean(percent, na.rm = TRUE))
```


## Eyes

1. ANOVA with factors MainGroup & Direction.

```{r eyes anova1, echo=FALSE, results = 'markup'}

# eyes ANCOVA 1
eyes_aov1 <- aov(eyes ~ maingroup * direction, data = aoi3_data)
summary(eyes_aov1)
# eyes_lsd1 <- LSD.test(eyes_aov1, "maingroup", group = FALSE)
# eyes_lsd1$comparison
```

2. ANOVA with factor MainGroup, for Forward only.

```{r eyes anova2, echo=FALSE, results = 'markup'}
# eyes ANOVA 2
eyes_aov2 <- aov(eyes ~ maingroup, data = filter(aoi3_data, direction == "forward"))
summary(eyes_aov2)
# eyes_lsd2 <- LSD.test(eyes_aov2, "maingroup", group = FALSE)
# eyes_lsd2$comparison
```

3. ANOVA with factor MainGroup, for Reverse only. 

```{r eyes anova3, echo=FALSE, results = 'markup'}
# eyes ANOVA 3
eyes_aov3 <- aov(eyes ~ maingroup, data = filter(aoi3_data, direction == "reversed"))
summary(eyes_aov3)
# eyes_lsd3 <- LSD.test(eyes_aov3, "maingroup", group = FALSE)
# eyes_lsd3$comparison
```

4. ANCOVA with factor Direction, and covariate AoASL and Age.

```{r eyes anova4, echo=FALSE, results = 'markup'}
# eyes ANOVA 4
eyes_aov4 <- aov(eyes ~ direction * aoasl * age, data = aoi3_data)
summary(eyes_aov4)
```

5. Regression with variables Age and AoASL, for Forward only

```{r eyes anova5, echo=FALSE, results = 'markup'}
# eyes ANOVA 5
eyes_aov5 <- lm(eyes ~ aoasl * age, data = filter(aoi3_data, direction == "forward"))
summary(eyes_aov5)
```

6. Regression with variables Age and AoASL, for Reverse only

```{r eyes anova6, echo=FALSE, results = 'markup'}
# eyes ANOVA 6
eyes_aov6 <- lm(eyes ~ aoasl * age, data = filter(aoi3_data, direction == "reversed"))
summary(eyes_aov6)
```

## Mouth

1. ANOVA with factors MainGroup & Direction.

```{r mouth anova1, echo=FALSE, results = 'markup'}

# mouth ANCOVA 1
mouth_aov1 <- aov(mouth ~ maingroup * direction, data = aoi3_data)
summary(mouth_aov1)
mouth_lsd1 <- LSD.test(mouth_aov1, "maingroup", group = FALSE)
mouth_lsd1$comparison
```

2. ANOVA with factor MainGroup, for Forward only.

```{r mouth anova2, echo=FALSE, results = 'markup'}
# mouth ANOVA 2
mouth_aov2 <- aov(mouth ~ maingroup, data = filter(aoi3_data, direction == "forward"))
summary(mouth_aov2)
# mouth_lsd2 <- LSD.test(mouth_aov2, "maingroup", group = FALSE)
# mouth_lsd2$comparison
```

3. ANOVA with factor MainGroup, for Reverse only. 

```{r mouth anova3, echo=FALSE, results = 'markup'}
# mouth ANOVA 3
mouth_aov3 <- aov(mouth ~ maingroup, data = filter(aoi3_data, direction == "reversed"))
summary(mouth_aov3)
# mouth_lsd3 <- LSD.test(mouth_aov3, "maingroup", group = FALSE)
# mouth_lsd3$comparison
```

4. ANCOVA with factor Direction, and covariate AoASL and Age.

```{r mouth anova4, echo=FALSE, results = 'markup'}
# mouth ANOVA 4
mouth_aov4 <- aov(mouth ~ direction * aoasl * age, data = aoi3_data)
summary(mouth_aov4)
```

5. Regression with variables Age and AoASL, for Forward only

```{r mouth anova5, echo=FALSE, results = 'markup'}
# mouth ANOVA 5
mouth_aov5 <- lm(mouth ~ aoasl * age, data = filter(aoi3_data, direction == "forward"))
summary(mouth_aov5)
```

6. Regression with variables Age and AoASL, for Reverse only

```{r mouth anova6, echo=FALSE, results = 'markup'}
# mouth ANOVA 6
mouth_aov6 <- lm(mouth ~ aoasl * age, data = filter(aoi3_data, direction == "reversed"))
summary(mouth_aov6)
```

## Neck

1. ANOVA with factors MainGroup & Direction.

```{r neck anova1, echo=FALSE, results = 'markup'}

# neck ANCOVA 1
neck_aov1 <- aov(neck ~ maingroup * direction, data = aoi3_data)
summary(neck_aov1)
# neck_lsd1 <- LSD.test(neck_aov1, "maingroup", group = FALSE)
# neck_lsd1$comparison
```

2. ANOVA with factor MainGroup, for Forward only.

```{r neck anova2, echo=FALSE, results = 'markup'}
# neck ANOVA 2
neck_aov2 <- aov(neck ~ maingroup, data = filter(aoi3_data, direction == "forward"))
summary(neck_aov2)
# neck_lsd2 <- LSD.test(neck_aov2, "maingroup", group = FALSE)
# neck_lsd2$comparison
```

3. ANOVA with factor MainGroup, for Reverse only. 

```{r neck anova3, echo=FALSE, results = 'markup'}
# neck ANOVA 3
neck_aov3 <- aov(neck ~ maingroup, data = filter(aoi3_data, direction == "reversed"))
summary(neck_aov3)
# neck_lsd3 <- LSD.test(neck_aov3, "maingroup", group = FALSE)
# neck_lsd3$comparison
```

4. ANCOVA with factor Direction, and covariate AoASL and Age.

```{r neck anova4, echo=FALSE, results = 'markup'}
# neck ANOVA 4
neck_aov4 <- aov(neck ~ direction * aoasl * age, data = aoi3_data)
summary(neck_aov4)
```

5. Regression with variables Age and AoASL, for Forward only

```{r neck anova5, echo=FALSE, results = 'markup'}
# neck ANOVA 5
neck_aov5 <- lm(neck ~ aoasl * age, data = filter(aoi3_data, direction == "forward"))
summary(neck_aov5)
```

6. Regression with variables Age and AoASL, for Reverse only

```{r neck anova6, echo=FALSE, results = 'markup'}
# neck ANOVA 6
neck_aov6 <- lm(neck ~ aoasl * age, data = filter(aoi3_data, direction == "reversed"))
summary(neck_aov6)
```

## FaceChest

We originally defined FaceChest such that

1. Face = eyes + mouth + chin
1. Chest = upperchest + midchest + lowerchest

BUT. Chin is actually neck. It's not even part of the face if you think about it. So I'm redefining FaceChest as:

1. Face = forehead + eyes + mouth
1. Chest = neck + upperchest + midchest + lowerchest

So let's do this. Then see what's happening across groups for FaceChest.

```{r calculate facechest, echo=FALSE, results = 'markup'}
# Calculate face, chest, and facechest - and moutheye too
fulldata <- fulldata %>%
  rowwise() %>%
  mutate(face = sum(forehead, eyes, mouth, na.rm = TRUE),
         chest = sum(neck, upperchest, midchest, lowerchest, na.rm = TRUE),
         facechest = (face - chest)/(face + chest),
         moutheye = (mouth - eyes)/(mouth + eyes))

# fulldata %>% 
#   gather(metric, value, c(facechest_old, facechest)) %>%
#   ggplot(aes(x = maingroup, y = value, fill = direction)) + geom_boxplot() + facet_wrap("metric")
```

Cool. Next we'll do error bar charts using the new FaceChest across groups. 

```{r facechest bars, message=FALSE, warning=FALSE}
facechest_info <- fulldata %>%
  filter(eye_exclude == FALSE) %>%
  group_by(maingroup, direction, participant) %>%
  summarise(facechest = mean(facechest, na.rm = TRUE)) %>%
  group_by(maingroup, direction) %>%
  summarise(mean = mean(facechest),
            sd = sd(facechest),
            n = n(),
            se = sd/sqrt(n))

ggplot(facechest_info, aes(x = maingroup, y = mean, fill = direction, color = direction)) +
  geom_point(stat = "identity", position = position_dodge(0.5), size = 2) + 
  geom_errorbar(aes(ymin = mean-se, ymax = mean+se), position = position_dodge(0.5), width = 0.3, size = 1) +
  labs(title = "FaceChest Ratio", subtitle = "Error bars represent SE", x = "", y = "facechest ratio") +
  scale_y_continuous(limits = c(-1,1)) + 
  geom_hline(yintercept = 0, linetype = "dotted")
```

Now let's do the ANOVAs. Also skipping LSDs here. 

1. ANOVA with factors MainGroup & Direction.

```{r facechest anova 1, echo=FALSE, results = 'markup', message=FALSE}
# Create the participant-level db (let's also pop moutheye in this too)
fc_data <- fulldata %>%
  filter(eye_exclude == FALSE) %>%
  select(id, maingroup, hearing, aoasl, age, direction, facechest, moutheye) %>%
  group_by(id, direction) %>%
  mutate(facechest = mean(facechest, na.rm = TRUE),
         moutheye = mean(moutheye, na.rm = TRUE)) %>%
  ungroup() %>%
  distinct()

# facechest ANCOVA 1
facechest_aov1 <- aov(facechest ~ maingroup * direction, data = fc_data)
summary(facechest_aov1)
# facechest_lsd1 <- LSD.test(facechest_aov1, "maingroup", group = FALSE)
# facechest_lsd1$comparison
```

2. ANOVA with factor MainGroup, for Forward only.

```{r facechest anova2, echo=FALSE, results = 'markup'}
# facechest ANOVA 2
facechest_aov2 <- aov(facechest ~ maingroup, data = filter(fc_data, direction == "forward"))
summary(facechest_aov2)
# facechest_lsd2 <- LSD.test(facechest_aov2, "maingroup", group = FALSE)
# facechest_lsd2$comparison
```

3. ANOVA with factor MainGroup, for Reverse only. 

```{r facechest anova3, echo=FALSE, results = 'markup'}
# facechest ANOVA 3
facechest_aov3 <- aov(facechest ~ maingroup, data = filter(fc_data, direction == "reversed"))
summary(facechest_aov3)
# facechest_lsd3 <- LSD.test(facechest_aov3, "maingroup", group = FALSE)
# facechest_lsd3$comparison
```

4. ANCOVA with factor Direction, and covariate AoASL and Age.

```{r facechest anova4, echo=FALSE, results = 'markup'}
# facechest ANOVA 4
facechest_aov4 <- aov(facechest ~ direction * aoasl * age, data = fc_data)
summary(facechest_aov4)
```

5. Regression with variables Age and AoASL, for Forward only

```{r facechest anova5, echo=FALSE, results = 'markup'}
# facechest ANOVA 5
facechest_aov5 <- lm(facechest ~ aoasl * age, data = filter(fc_data, direction == "forward"))
summary(facechest_aov5)
```

6. Regression with variables Age and AoASL, for Reverse only

```{r facechest anova6, echo=FALSE, results = 'markup'}
# facechest ANOVA 6
facechest_aov6 <- lm(facechest ~ aoasl * age, data = filter(fc_data, direction == "reversed"))
summary(facechest_aov6)
```

# Eye Gaze & Performance Correlations

Next we're going to correlate our eye gaze metrics (Eye, Mouth, Neck, and FaceChest) with lexical recall and gist. Okay! But remember we have a slightly smaller dataset here because we've excluded some participants for having bad eye data (but they had valid behavioral data so we kept them in the AoA-Performance correlations above). 

**We also removed gist_forward column.**

```{r repackage data for correlations again}
eyeperf <- fulldata %>%
  filter(eye_exclude == FALSE) %>%
  select(participant, maingroup, hearing, direction, acc, gist, eyes, mouth, neck, facechest) %>%
  group_by(maingroup, participant, direction) %>%
  mutate(gist = mean(gist, na.rm = TRUE),
         lex = mean(acc, na.rm = TRUE),
         eyes = mean(eyes, na.rm = TRUE),
         mouth = mean(mouth, na.rm = TRUE),
         neck = mean(neck, na.rm = TRUE),
         facechest = mean(facechest, na.rm = TRUE)) %>%
  ungroup() %>%
  select(maingroup, participant, hearing, direction, gist, lex, eyes, mouth, neck, facechest) %>%
  distinct() %>%
  gather(metric, value, gist:facechest) %>%
  unite(metricvalue, c(metric, direction), sep = "_") %>%
  spread(metricvalue, value) %>%
  select(-participant, -maingroup) %>%
  select(hearing, gist_reversed, lex_forward, lex_reversed, eyes_forward, eyes_reversed,
         mouth_forward, mouth_reversed, neck_forward, neck_reversed, facechest_forward, facechest_reversed)

eyeperf_deaf <- eyeperf %>% filter(hearing == "Deaf") %>% select(-hearing)
eyeperf_hearing <- eyeperf %>% filter(hearing == "Hearing") %>% select(-hearing)
eyeperf_all <- eyeperf %>% select(-hearing)


# Correlations for Deaf
print("DEAF Correlations - Pearson's r")
#corstarsl(lexgist_deaf)
Hmisc::rcorr(as.matrix(eyeperf_deaf))$r
print("DEAF Correlations - P-values")
Hmisc::rcorr(as.matrix(eyeperf_deaf))$P
cat(paste("","\n",""))

# Correlations for Hearing
print("HEARING Correlations - Pearson's r")
#corstarsl(lexgist_hearing)
Hmisc::rcorr(as.matrix(eyeperf_hearing))$r
print("HEARING Correlations - P-values")
Hmisc::rcorr(as.matrix(eyeperf_hearing))$P
cat(paste("","\n",""))

# Correlations for All
print("ALL Correlations - Pearson's r")
#corstarsl(lexgist_all)
Hmisc::rcorr(as.matrix(eyeperf_all))$r
print("ALL Correlations - P-values")
Hmisc::rcorr(as.matrix(eyeperf_all))$P

```

I'm also including nicely formatted tables with *** indicators of significance for quick referencing. Order: Deaf, Hearing, All. 

```{r pretty correlations for eyeperf}
corstarsl(eyeperf_deaf)
corstarsl(eyeperf_hearing)
corstarsl(eyeperf_all)
```

And the correlation table.

```{r eyeperf correlation charts, fig.height=12, fig.width=12, message=FALSE, warning=FALSE}
ggpairs(eyeperf, columns = c(2:11), aes(color = hearing))
```

# Heat Maps
And finally, we're going to do heat maps. 
```{r heat map}
eyegaze_heat <- fulldata %>%
  ungroup() %>%
  filter(eye_exclude == FALSE) %>%
  select(id:direction, belly, lowerchest, midchest, upperchest, neck, mouth, eyes, forehead) %>%
  gather(aoi, percent, belly:forehead) %>%
  group_by(maingroup, participant, direction, aoi) %>%
  summarise(percent = mean(percent, na.rm=TRUE)) %>%
  group_by(maingroup,direction,aoi) %>%
  summarise(percent = mean(percent, na.rm=TRUE)) %>%
  ungroup() %>%
  filter(!is.na(aoi)) %>%
  mutate(aoi = factor(aoi,levels=c("belly","lowerchest","midchest",
                                   "upperchest","neck","mouth","eyes","forehead")))

eyegaze_heat_all <- fulldata %>%
  ungroup() %>%
  filter(eye_exclude == FALSE) %>%
  select(id:direction, belly, lowerchest, midchest, upperchest, neck, mouth, eyes, forehead) %>%
  gather(aoi, percent, belly:forehead) %>%
  group_by(maingroup,participant,direction,aoi) %>%
  dplyr::summarize(percent = mean(percent, na.rm=TRUE)) %>%
  group_by(maingroup,direction,aoi) %>%
  dplyr::summarize(percent = mean(percent, na.rm=TRUE)) %>%
  group_by(maingroup,aoi) %>%
  dplyr::summarize(percent = mean(percent, na.rm=TRUE)) %>%
  ungroup() %>%
  filter(!is.na(aoi)) %>%
  mutate(aoi = factor(aoi,levels=c("belly","lowerchest","midchest",
                                   "upperchest","neck","mouth","eyes","forehead")))

ggplot(eyegaze_heat, aes(x = maingroup, y = aoi)) +
  geom_tile(aes(fill=percent),color="lightgray",na.rm=TRUE) + 
#  scale_fill_gradient(low = "lightblue",high = "steelblue") +
#  scale_fill_distiller(type="div", palette = "RdYlBu") +
  scale_fill_viridis(option = "viridis", direction=-1, limits = c(0,.75)) +
  theme(axis.text.x=element_text(angle=45,hjust=1)) + facet_grid(. ~ direction) +
  ylab("") + xlab("") + ggtitle("Eye Gaze Heat Map, by Direction")

ggplot(eyegaze_heat_all, aes(x = maingroup, y = aoi)) +
  geom_tile(aes(fill=percent),color="lightgray",na.rm=TRUE) + 
#  scale_fill_gradient(low = "lightblue",high = "steelblue") +
#  scale_fill_distiller(type="div", palette = "RdYlBu") +
  scale_fill_viridis(option = "viridis", direction=-1, limits = c(0,.75)) +
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  ylab("") + xlab("") + ggtitle("Eye Gaze Heat Map (Direction Collapsed)")
```

# Summary

Below are the p-values from the ANOVAs with 4 MainGroups. I never included Age as a covariate because it never improved the model. I included all ANOVAs for Gist and Lex Recall, and ANOVAs for any eye AOI or ratio was included only if either maingroup or direction was significant. Deafearly-Deaflate shows the LSD p-value for that comparison.
```{r results1, rows.print = 20}
results1 <- structure(list(model = c("gist-maingroup-both", "gist-maingroup-fw", 
"gist-maingroup-rv", "lexrecall-maingroup-both", "lexrecall-maingroup-fw", 
"lexrecall-maingroup-rv", "mouth-maingroup-both", "upperchest-maingroup-both", 
"upperchest-maingroup-rv", "facechest-maingroup-both", "moutheye-maingroup-both"
), maingroup = c(0, 0, 0.01, 0, 0.04, 0.02, 0.06, 0, 0.01, 0.1, 
0.05), direction = c(0, NA, NA, 0, NA, NA, 0.06, 0.16, NA, 0.07, 
0.48), `deafearly-deaflate` = c(0.1, 0.69, 0.02, 0.11, 0.95, 
0.06, 0.38, 0.94, 0.52, 0.08, 0.68)), .Names = c("model", "maingroup", 
"direction", "deafearly-deaflate"), class = c("tbl_df", "tbl", 
"data.frame"), row.names = c(NA, -11L))

results1
```

And below are the p-values from the ANCOVAs with Hearing & AoASL. I included all ANCOVAs for Gist and Lex Recall, and ANCOVAs for any eye AOI or ratio was included only if any main factor was significant. LSD comparisons are not needed because there's only 2 levels in each group!

```{r results2, rows.print = 25}
results2 <- structure(list(model = c("gist-both", "gist-fw", "gist-rv", "lex-both", 
"lex-fw", "lex-rv", "forehead-fw", "mouth-both", "mouth-rv", 
"upperchest-both", "upperchest-rv", "facechest-both", "moutheye-both"
), hearing = c(0, 0.00, 0.01, 0.01, 0.22, 0.03, 0.06, 0.01, 
0.04, 0.01, 0.01, 0.35, 0.07), direction = c(0, NA, NA, 0, NA, 
NA, NA, 0.05, NA, 0.21, NA, 0.05, 0.52), aoasl = c(0.22, 0.77, 
0.19, 0.56, 0.58, 0.25, 0.08, 0.06, 0.12, 0.68, 0.95, 0.12, 0.44
), age = c(0.08, 0.01, 0.86, 0.09, 0.02, 0.7, 0.68, 0.28, 0.5, 
0.02, 0.08, 0.00, 0.21)), .Names = c("model", "hearing", "direction", 
"aoasl", "age"), class = c("tbl_df", "tbl", "data.frame"), row.names = c(NA, 
-13L))
results2
```

Finally, the correlations for Deaf and Hearing separately are not significant. But there are significant correlations across all participants. I worry it is caused by HearingNovice, though...

```{r correlation table}
results3 <- tribble(
  ~ metric, ~ AoASLcorrelationRvalue, ~ Pvalue,
  "gist-fw", -0.32, 0.019,
  "gist-rv", -0.39, 0.004,
  "lex-fw", -0.08, 0.567,
  "lex-rv", -0.34, 0.014
)
results3
```

# Ternary Plots

Let's make triangle plots. "What?" you say. Read on. 

```{r fig.height=12, fig.width=12, message=FALSE, warning=FALSE}
library(ggtern)
fulldata %>% 
  ggtern(aes(x = eyes, y = mouth, z = neck)) + facet_grid(direction ~ maingroup) + stat_density_tern(geom='polygon', aes(fill=..level..), bins=4) + geom_point(color = "white", alpha = 0.5) + theme_bw()

fulldata %>% 
  ggtern(aes(x = eyes, y = mouth, z = neck)) + facet_grid(direction ~ maingroup) + geom_confidence_tern(breaks = c(.5), color = "red") + geom_point() + theme_bw()
```

# Rain's Notes
About Adults:

-I think I want to write it up as an ANCOVA, with direction included.  And LSD comparisons instead of Tukey.  (I will do my own corrections)
-You often have one liners summarizing results, in all tabs, those are nice, keep them coming.
-(If you have reasons to present anything other than the ANCOVA, put that in your results tab)
 
I think if we do it this way then we get a really important story to tell:  That the *critical* AoA cutoff is below 4 vs above 4 years of age (two groups 0-4 vs 4-13).  This suggest early ASL is important.  

```{r eval=FALSE, include=FALSE}
# Gist as binomial logit-link function model
gist_glmm <- glmer(gist ~ direction * maingroup + (1|id) + (1|story), data = fulldata, family=binomial (link="logit"))
summary(gist_glmm)
```
