---
title: "The Last Data Analysis to Rule Them All? (study1adults)"
author: "Adam Stone, PhD" 
date: '`r format(Sys.Date(), "%m-%d-%Y")`'
output:
  github_document:
    toc: yes
    toc_depth: 2
  html_notebook:
    code_folding: hide
    theme: paper
    highlight: tango
    toc: yes
    toc_depth: 2
    toc_float: yes
    df_print: paged
---

# Putting It All Back Together

Throughout all the data analysis we've done, the datasets have become more fragmented - lexical recall, gist, and eye tracking datasets. I want to put them all together in one whole dataset again so we can perform some analyses more efficiently (particularly correlations). The only thing I need to remember is we'll have a new column called `eye_exclude` and if it is set to `TRUE` it means we can't include that row in any analysis relating to eye gaze (usually because that trial was less than 25% looking). 

```{r smash data together, message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(stringr)
library(lme4)
library(lmerTest)
library(scales)

# Load lex and eye data
cleanlexdata <- read_csv("cleandata.csv") %>%
  select(-(forehead:total))

cleaneyedata <- read_csv("cleanpercentdata.csv") %>%
  spread(aoi,percent) %>%
  add_column(eye_exclude = FALSE)

# What rows were removed from the eye data back in 03eyegaze? Let's add back in
# With a new column - eye_exclude
removed <- anti_join(cleanlexdata, cleaneyedata) %>%
  add_column(eye_exclude = TRUE)
eyelexdata <- bind_rows(cleaneyedata, removed)

# Load gist data
gist <- read_csv('gist_indiv.csv', col_types = cols(
  participant = col_character(),
  gist.fw1 = col_integer(),
  gist.rv2 = col_integer(),
  gist.fw3 = col_integer(),
  gist.rv4 = col_integer()
)) %>%
  gather(video, gist, gist.fw1:gist.rv4) %>%
  mutate(video = str_sub(video,6,8))

# Presto, our full reunified dataset - 'fulldata'
fulldata <- left_join(eyelexdata, gist)
```

# Group Changes and Participant Tables

We have some changes to make to the groups. First, fix Josh as learning ASL when he was 6. Next, drop the DeafNative Group and reclassify all who learned ASL < 3.9 as DeafEarly and ASL => 4.0 as DeafLate. 

```{r groups and participant tables}
# Change Josh's AoASL to 6
fulldata <- fulldata %>%
  mutate(aoasl = as.double(aoasl)) %>%
  mutate(aoasl = case_when(
    participant == "Josh" ~ 6,
    TRUE ~ aoasl
  ))

# Reclassify Groups
fulldata <- fulldata %>%
  mutate(maingroup = case_when(
    hearing == "Deaf" & aoasl < 4 ~ "DeafEarly",
    hearing == "Deaf" & aoasl >= 4 ~ "DeafLate",
    maingroup == "HearingLateASL" ~ "HearingLate",
    maingroup == "HearingNoviceASL" ~ "HearingNovice"
  ))

# Create Participant Demographics Table
participant_info <- fulldata %>%
  select(-(acc:gist)) %>%
  select(-(video:direction)) %>%
  distinct() %>% 
  group_by(maingroup) %>%
  summarise(n = n(),
            age_mean = mean(age),
            age_sd = sd(age),
            aoasl_mean = mean(aoasl),
            aoasl_sd = sd(aoasl),
            signyrs_mean = mean(signyrs),
            signyrs_sd = sd(signyrs),
            selfrate_mean = mean(selfrate),
            selfrate_sd = sd(selfrate)) %>%
  ungroup() %>%
  mutate_if(is.double, funs(round(., 2))) %>%
  mutate(age = paste(age_mean, "±", age_sd, sep = " "),
         aoasl = paste(aoasl_mean, "±", aoasl_sd, sep = " "),
         signyrs = paste(signyrs_mean, "±", signyrs_sd, sep = " "),
         selfrate = paste(selfrate_mean, "±", selfrate_sd, sep = " ")) %>%
  select(-(age_mean:selfrate_sd)) %>%
  print()
```

# Gist & Lexical Recall Data

## Tables & Charts

Let's generate a table for lexical recall and gist for forward vs. reversed stories. 

```{r lex & gist table}
lexgist_info <- fulldata %>%
  group_by(maingroup, direction) %>%
  summarise(lex_mean = mean(acc, na.rm = TRUE),
            lex_sd = sd(acc, na.rm = TRUE),
            gist_mean = mean(gist),
            gist_sd = sd(gist)) %>%
  ungroup() %>%
  mutate_if(is.double, funs(round(., 2))) %>% 
  mutate(lex = paste(lex_mean, "±", lex_sd, sep = " "),
         gist = paste(gist_mean, "±", gist_sd, sep = " ")) %>%
  select(-(lex_mean:gist_sd)) %>%
  gather(metric, value, lex:gist) %>%
  unite("metric", c(metric, direction), sep = "_") %>%
  spread(metric, value) %>%
  print()
```

And then bar charts too after that with error bars. 

```{r lex and gist bar charts}
# Gist bar chart
gist_bar <- fulldata %>% select(participant, maingroup, direction, gist) %>%
  group_by(maingroup, participant, direction) %>%
  summarise(gist = mean(gist)) %>%
  group_by(maingroup, direction) %>%
  summarise(mean = mean(gist),
            sd = sd(gist),
            count = n(),
            se = sd/sqrt(count))

ggplot(gist_bar, aes(x = maingroup, y = mean, fill = direction)) +
  geom_bar(stat = "identity", position = position_dodge()) + 
  geom_errorbar(aes(ymin = mean-se, ymax = mean+se), position = position_dodge(0.9), width = 0.5) +
  labs(title = "Story Comprehension (Gist)", subtitle = "Error bars represent SE", x = "", y = "mean gist") +
  scale_y_continuous(labels = percent)

# Gist bar chart
lex_bar <- fulldata %>% select(participant, maingroup, direction, acc) %>%
  group_by(maingroup, participant, direction) %>%
  summarise(acc = mean(acc, na.rm = TRUE)) %>%
  group_by(maingroup, direction) %>%
  summarise(mean = mean(acc, na.rm = TRUE),
            sd = sd(acc, na.rm = TRUE),
            count = n(),
            se = sd/sqrt(count))

ggplot(lex_bar, aes(x = maingroup, y = mean, fill = direction)) +
  geom_bar(stat = "identity", position = position_dodge()) + 
  geom_errorbar(aes(ymin = mean-se, ymax = mean+se), position = position_dodge(0.9), width = 0.5) +
  labs(title = "Lexical Recall", subtitle = "Error bars represent SE", x = "", y = "mean accuracy") +
  scale_y_continuous(labels = percent, limits = c(0,1))
```

## ANOVAs

Here, ANOVAs are done with **Direction Included.** Gist first, with Tukey's comparisons. 

```{r gist anova}
# Gist ANOVA. We have to make participant-level means first. 
gist_aov <- fulldata %>%
  group_by(maingroup, participant, direction) %>%
  summarise(gist = mean(gist, na.rm = TRUE)) %>%
#  filter(direction != "reversed") %>%
  lm(gist ~ maingroup * direction, data = .)
summary(aov(gist_aov))
TukeyHSD(aov(gist_aov))
```

And Lexical Recall ANOVA with Tukey's comparisons. 

```{r lex anova}
# Lex ANOVA. We have to make participant-level means first. 
lex_aov <- fulldata %>%
  group_by(maingroup, participant, direction) %>%
  summarise(acc = mean(acc, na.rm = TRUE)) %>%
#  filter(direction != "reversed") %>%
  lm(acc ~ maingroup * direction, data = .)
summary(aov(lex_aov))
TukeyHSD(aov(lex_aov))
```

## AoA Correlations

Next, we want to look at correlations between AoA and Gist, and betwen AoA and Lexical Recall. Rain asked for forward and reversed separately (1) deaf only, (2) hearing only, and (3) both. Let's make it work. 

```{r repackage data for correlations, message=FALSE, warning=FALSE}
# Let's make participant-level data, and have forward/reversed in separate columns
lexgist_data <- fulldata %>%
  group_by(maingroup, participant, direction) %>%
  mutate(gist = mean(gist, na.rm = TRUE),
         lex = mean(acc, na.rm = TRUE)) %>%
  ungroup() %>%
  select(maingroup, participant, hearing, direction, aoasl, gist, lex) %>%
  distinct() %>%
  gather(metric, value, gist:lex) %>%
  unite(metricvalue, c(metric, direction), sep = "_") %>%
  spread(metricvalue, value) %>%
  select(-participant, -maingroup)

lexgist_deaf <- lexgist_data %>% filter(hearing == "Deaf") %>% select(-hearing)
lexgist_hearing <- lexgist_data %>% filter(hearing == "Hearing") %>% select(-hearing)
lexgist_all <- lexgist_data %>% select(-hearing)

# Load awesome function to make correlation tables with stars for significance
# From: https://myowelt.blogspot.co.uk/2008/04/beautiful-correlation-tables-in-r.html
corstarsl <- function(x){ 
require(Hmisc) 
x <- as.matrix(x) 
R <- Hmisc::rcorr(x)$r 
p <- Hmisc::rcorr(x)$P 
## define notions for significance levels; spacing is important.
mystars <- ifelse(p < .001, "***", ifelse(p < .01, "** ", ifelse(p < .05, "* ", " ")))
## trunctuate the matrix that holds the correlations to two decimal
R <- format(round(cbind(rep(-1.11, ncol(x)), R), 2))[,-1] 
## build a new matrix that includes the correlations with their apropriate stars 
Rnew <- matrix(paste(R, mystars, sep=""), ncol=ncol(x)) 
diag(Rnew) <- paste(diag(R), " ", sep="") 
rownames(Rnew) <- colnames(x) 
colnames(Rnew) <- paste(colnames(x), "", sep="") 
## remove upper triangle
Rnew <- as.matrix(Rnew)
Rnew[upper.tri(Rnew, diag = TRUE)] <- ""
Rnew <- as.data.frame(Rnew) 
## remove last column and return the matrix (which is now a data frame)
Rnew <- cbind(Rnew[1:length(Rnew)-1])
return(Rnew) 
}

# Correlations for Deaf
print("DEAF Correlations - Pearson's r")
#corstarsl(lexgist_deaf)
Hmisc::rcorr(as.matrix(lexgist_deaf))$r
print("DEAF Correlations - P-values")
Hmisc::rcorr(as.matrix(lexgist_deaf))$P
cat(paste("","\n",""))

# Correlations for Hearing
print("HEARING Correlations - Pearson's r")
#corstarsl(lexgist_hearing)
Hmisc::rcorr(as.matrix(lexgist_hearing))$r
print("HEARING Correlations - P-values")
Hmisc::rcorr(as.matrix(lexgist_hearing))$P
cat(paste("","\n",""))

# Correlations for All
print("ALL Correlations - Pearson's r")
#corstarsl(lexgist_all)
Hmisc::rcorr(as.matrix(lexgist_all))$r
print("ALL Correlations - P-values")
Hmisc::rcorr(as.matrix(lexgist_all))$P

```

I'm also including nicely formatted tables with *** indicators of significance for quick referencing. Order: Deaf, Hearing, All. 

```{r pretty correlations}
corstarsl(lexgist_deaf)
corstarsl(lexgist_hearing)
corstarsl(lexgist_all)
```

# Eye Gaze Data

Now eye gaze data. Boxplots first. Also here, we're renaming "chin" to "neck" because that's what it actually is! 
```{r eye gaze boxplot}
# rename chin to neck
fulldata <- fulldata %>%
  rename(neck = chin)

fulldata %>%
  filter(eye_exclude == FALSE) %>%
  select(direction, moutheye:upperchest) %>%
  select(-moutheye, -facechest, -face, -chest) %>%
  gather(aoi, percent, belly:upperchest) %>%
  ggplot(aes(x = aoi, y = percent, fill = direction)) + geom_boxplot()
```

## ANOVAS
We have three big AOIs. Eyes, Mouth, and Neck. To some extent we also have Forehead and UpperChest. Let's run ANOVAs on all those five AOIs and see if either direction or group membership affect looking behavior for any of those AOIs. 

```{r}
# Define the 5 AOIs
aoi5 <- c('forehead','eyes','mouth','neck','upperchest')

# Organize the data by those AOIs (but again we need participant-level first)
aoi5_data <- fulldata %>%
  filter(eye_exclude == FALSE) %>%
  select(id, maingroup, direction, belly:upperchest) %>%
  gather(aoi, percent, belly:upperchest) %>%
  group_by(id, direction, aoi) %>%
  mutate(percent = mean(percent, na.rm = TRUE)) %>%
  ungroup () %>%
  distinct() %>%
  filter(aoi %in% aoi5) %>%
  spread(aoi, percent)

print("Forehead ANOVA")
forehead_aov <- lm(forehead ~ maingroup * direction, data = aoi5_data)
summary(aov(forehead_aov))
cat(paste("","\n",""))

print("Eyes ANOVA")
eyes_aov <- lm(eyes ~ maingroup * direction, data = aoi5_data)
summary(aov(eyes_aov))
cat(paste("","\n",""))

print("Mouth ANOVA")
mouth_aov <- lm(mouth ~ maingroup * direction, data = aoi5_data)
summary(aov(mouth_aov))
cat(paste("","\n",""))

print("Neck ANOVA")
neck_aov <- lm(neck ~ maingroup * direction, data = aoi5_data)
summary(aov(neck_aov))
cat(paste("","\n",""))

print("UpperChest ANOVA")
upperchest_aov <- lm(upperchest ~ maingroup * direction, data = aoi5_data)
summary(aov(upperchest_aov))
cat(paste("","\n",""))
```

## FaceChest

# Rain's Notes
About Adults:

-I think I want to write it up as an ANCOVA, with direction included.  And LSD comparisons instead of Tukey.  (I will do my own corrections)
-You often have one liners summarizing results, in all tabs, those are nice, keep them coming.
-(If you have reasons to present anything other than the ANCOVA, put that in your results tab)
 
I think if we do it this way then we get a really important story to tell:  That the *critical* AoA cutoff is below 4 vs above 4 years of age (two groups 0-4 vs 4-13).  This suggest early ASL is important.  
 
