---
title: "ANOVAs Grwarrrr"
output:
  html_notebook:
    toc: yes
    toc_depth: 2
    toc_float: yes
---

First I'm loading the behavioraldata.csv file. And make sure all factor columns are identified as factors and all numeric columns are identified as numeric.

```{r message=FALSE, warning=TRUE}
library(tidyverse) 
library(afex) 
bdata <- read_csv("behavioraldata.csv") %>%
  mutate(id = factor(id),
         participant = factor(participant),
         hearing = factor(hearing),
         maingroup = factor(maingroup),
         direction = factor(direction))
bdata
```

# Gist

Cool. Now I'm running a 2x4 ANOVA on Gist.

* Outcome: gist
* Within subjects: direction
* Between subjects: maingroup 

```{r message=FALSE}
a1 <- aov_ez("id", "gist", bdata, between = "maingroup", within = "direction")
a1
```

Next, I run a Forward ANOVA on Gist (between subjects: maingroup), with least-square means posthoc analyses for all pairs, with **no p-value corrections.**

```{r message=FALSE}
a2 <- aov_ez("id", "gist", filter(bdata, direction == "forward"), between = "maingroup")
a2
lsmeans(a2, "maingroup", contr = "pairwise", adjust = "none")$contrasts
```

Next, I run a Reverse ANOVA on Gist (between subjects: maingroup), with least-square means posthoc analyses for all pairs, with **no p-value corrections.**

```{r message=FALSE}
a3 <- aov_ez("id", "gist", filter(bdata, direction == "reversed"), between = "maingroup")
a3
lsmeans(a3, "maingroup", contr = "pairwise", adjust = "none")$contrasts
```

# Lexical Recall

Next is a 2x4 ANOVA on Lexical Recall.

* Outcome: lexical recall
* Within subjects: direction
* Between subjects: maingroup 

```{r message=FALSE}
a4 <- aov_ez("id", "acc", bdata, between = "maingroup", within = "direction")
a4
```

Because there is no significant interaction between Subject Group and Direction, I'm going to do posthoc analyses on this anova, averaging across direction. 

```{r}
lsmeans(a4, "maingroup", contr = "pairwise", adjust = "none")$contrasts
```


```{r eval=FALSE, message=FALSE, include=FALSE}
# Next, I run a Forward ANOVA on Lex Recall (between subjects: maingroup), with least-square means posthoc analyses for all pairs, with **no p-value corrections.**

a5 <- aov_ez("id", "acc", filter(bdata, direction == "forward"), between = "maingroup")
a5
lsmeans(a5, "maingroup", contr = "pairwise", adjust = "none")$contrasts

#Next, I run a Reverse ANOVA on Lex Recall (between subjects: maingroup), with least-square means posthoc analyses for all pairs, with **no p-value corrections.**

a6 <- aov_ez("id", "acc", filter(bdata, direction == "reversed"), between = "maingroup")
a6
lsmeans(a6, "maingroup", contr = "pairwise", adjust = "none")$contrasts
```

Let's do a mu test for reversed videos to make sure accuracy is significantly different than 50%. 

```{r}
rev <- filter(bdata, direction == "reversed") 
rev <- unlist(rev$acc)
t.test(rev, mu=0.5)
```


# Eye Gaze
Next I load the aoidata.csv file. 

```{r message=FALSE, warning=TRUE}
edata <- read_csv("aoidata.csv") %>%
  mutate(id = factor(id),
         hearing = factor(hearing),
         maingroup = factor(maingroup),
         direction = factor(direction),
         aoi = factor(aoi))
edata
```

Next we run a 3x2x4 ANOVA on eye gaze. 

* Outcome: percent looking
* Within subjects: direction & AOI
* Between subjects: maingroup

```{r message=FALSE}
e1 <- aov_ez("id", "percent", edata, between = "maingroup", within = c("aoi","direction"))
e1
summary(e1)

lsmeans(e1, "maingroup", contr = "pairwise", adjust = "none")$contrasts

```

Then again for Eyes only but with only one factor: video direction. 

```{r}
e2 <- aov_ez("id", "percent", filter(edata, aoi == "eyes"), within = "direction")
e2
```

Mouth only but with only one factor: video direction. 

```{r}
e3 <- aov_ez("id", "percent", filter(edata, aoi == "mouth"), within = "direction")
e3
```

Neck only but with only one factor: video direction. 

```{r}
e4 <- aov_ez("id", "percent", filter(edata, aoi == "neck"), within = "direction")
e4
```

# High vs. Low Performers
First I run 09finaldata and get the `fulldata` tibble to change it to `eyeperf`
```{r}
eyeperf <- fulldata %>%
  filter(eye_exclude == FALSE) %>%
  select(participant, maingroup, hearing, direction, aoasl, signyrs, selfrate, acc, gist, eyes, mouth, neck, facechest) %>%
  group_by(maingroup, participant, direction) %>%
  mutate(gist = mean(gist, na.rm = TRUE),
         lex = mean(acc, na.rm = TRUE),
         eyes = mean(eyes, na.rm = TRUE),
         mouth = mean(mouth, na.rm = TRUE),
         neck = mean(neck, na.rm = TRUE),
         facechest = mean(facechest, na.rm = TRUE)) %>%
  ungroup() %>%
  select(maingroup, participant, hearing, direction, aoasl, signyrs, selfrate, gist, lex, eyes, mouth, neck, facechest) %>%
  distinct() %>%
  gather(metric, value, gist:facechest) %>%
  unite(metricvalue, c(metric, direction), sep = "_") %>%
  spread(metricvalue, value) %>%
  select(-participant) %>%
  select(maingroup, hearing, aoasl, signyrs, selfrate, gist_reversed, lex_forward, lex_reversed, eyes_forward,
         eyes_reversed, mouth_forward, mouth_reversed, neck_forward, neck_reversed,
         facechest_forward, facechest_reversed)

# eyeperf <- fulldata %>%
#   filter(eye_exclude == FALSE) %>%
#   select(participant, maingroup, hearing, direction, aoasl, signyrs, selfrate, acc, gist, upperchest, midchest, lowerchest, belly) %>%
#   group_by(maingroup, participant, direction) %>%
#   mutate(gist = mean(gist, na.rm = TRUE),
#          lex = mean(acc, na.rm = TRUE),
#          upperchest = mean(upperchest, na.rm = TRUE),
#          midchest = mean(midchest, na.rm = TRUE),
#          lowerchest = mean(lowerchest, na.rm = TRUE),
#          belly = mean(belly, na.rm = TRUE)) %>%
#   ungroup() %>%
#   select(maingroup, participant, hearing, direction, aoasl, signyrs, selfrate, gist, lex, upperchest, midchest, lowerchest, belly) %>%
#   distinct() %>%
#   gather(metric, value, gist:belly) %>%
#   unite(metricvalue, c(metric, direction), sep = "_") %>%
#   spread(metricvalue, value) %>%
#   select(-participant, -maingroup)

# eyeperf <- fulldata %>%
#   filter(eye_exclude == FALSE) %>%
#   select(participant, maingroup, story, hearing, direction, aoasl, signyrs, selfrate, acc, gist, upperchest, midchest, lowerchest, belly) %>%
#   group_by(maingroup, participant, direction, story) %>%
#   mutate(chest = sum(upperchest, midchest, lowerchest, belly, na.rm = TRUE)) %>%
#   group_by(maingroup, participant, direction) %>%
#   mutate(gist = mean(gist, na.rm = TRUE),
#          lex = mean(acc, na.rm = TRUE),
#          chest = mean(chest, na.rm = TRUE)) %>%
#   ungroup() %>%
#   select(maingroup, participant, hearing, direction, aoasl, signyrs, selfrate, gist, lex, chest) %>%
#   distinct() %>%
#   gather(metric, value, gist:chest) %>%
#   unite(metricvalue, c(metric, direction), sep = "_") %>%
#   spread(metricvalue, value) %>%
#   select(-participant, -maingroup)
```

## Gist Reversed
Cool. Now let's divide into high and low performers. I'll  do gist reversed. Well, maybe that won't work because that's only 3 levels! So I'll combine 0.5 and 1 (that's 32) vs 0 (19). 

```{r}
# gist_hl <- eyeperf %>% 
#   select(gist_reversed, facechest_reversed) %>%
#   mutate(group = case_when(
#     gist_reversed == 0 ~ "low",
#     gist_reversed == 0.5 ~ "high",
#     gist_reversed == 1 ~ "high"
#   )) %>%
#   mutate(group = factor(group)) %>%
#   select(-gist_reversed) 
# 
# gist_high = filter(gist_hl, group == "high") %>% select(-group)
# gist_low = filter(gist_hl, group == "low") %>% select(-group)
# 
# gist_high = unlist(gist_high$facechest_reversed)
# gist_low = unlist(gist_low$facechest_reversed)
# t.test(gist_high, gist_low)


gist_hl <- eyeperf %>% 
  select(gist_reversed, facechest_reversed) %>%
  mutate(group = case_when(
    gist_reversed == 0 ~ "low",
    gist_reversed == 0.5 ~ "medium",
    gist_reversed == 1 ~ "high"
  )) %>%
  mutate(group = factor(group)) %>%
  select(-gist_reversed) 

gist_m <- aov(data = gist_hl, facechest_reversed ~ group)
summary(gist_m)

gist_hl %>% group_by(group) %>% summarise(fcr = mean(facechest_reversed, na.rm = TRUE))
```

## Lex Recall Forward (Mean)
p = 0.421 (doesn't matter if the mean score is in high or low group)
```{r}
lex_fw_hl <- eyeperf %>% 
  select(lex_forward, chest_forward) 
lex_mean = mean(lex_fw_hl$lex_forward)

lex_fw_hl <- lex_fw_hl %>%
  mutate(group = case_when(
    lex_forward > lex_mean ~ "high",
    lex_forward <= lex_mean ~ "low")) %>%
  mutate(group = factor(group))

summary(lex_fw_hl)

lex_fw_high = filter(lex_fw_hl, group == "high") %>% select(chest_forward)
lex_fw_low = filter(lex_fw_hl, group == "low") %>% select(chest_forward)

lexfw_high = unlist(lex_fw_high$chest_forward)
lexfw_low = unlist(lex_fw_low$chest_forward)
t.test(lexfw_high, lexfw_low)
```

## Lex Recall Forward (Median)
This is interesting, if high is > median and low is <= median, the p-value is 0.066. If high is >= median and low is < median, the p-value is 0.24. 
```{r}
lex_fw_hl <- eyeperf %>% 
  select(lex_forward, chest_forward) 
lex_med = median(lex_fw_hl$lex_forward)

lex_fw_hl <- lex_fw_hl %>%
  mutate(group = case_when(
    lex_forward > lex_med ~ "high",
    lex_forward <= lex_med ~ "low")) %>%
  mutate(group = factor(group))

summary(lex_fw_hl)

lex_fw_high = filter(lex_fw_hl, group == "high") %>% select(chest_forward)
lex_fw_low = filter(lex_fw_hl, group == "low") %>% select(chest_forward)

lexfw_high = unlist(lex_fw_high$chest_forward)
lexfw_low = unlist(lex_fw_low$chest_forward)
t.test(lexfw_high, lexfw_low)
```

## Lex Recall Reverse (Mean)
p = 0.8973 (doesn't matter if the mean score is in high or low group)
```{r}
lex_fw_hl <- eyeperf %>% 
  select(lex_reversed, chest_reversed) 
lex_mean = mean(lex_fw_hl$lex_reversed, na.rm = TRUE)

lex_fw_hl <- lex_fw_hl %>%
  mutate(group = case_when(
    lex_reversed > lex_mean ~ "high",
    lex_reversed <= lex_mean ~ "low")) %>%
  mutate(group = factor(group))

summary(lex_fw_hl)

lex_fw_high = filter(lex_fw_hl, group == "high") %>% select(chest_reversed)
lex_fw_low = filter(lex_fw_hl, group == "low") %>% select(chest_reversed)

lexfw_high = unlist(lex_fw_high$chest_reversed)
lexfw_low = unlist(lex_fw_low$chest_reversed)
t.test(lexfw_high, lexfw_low)
```

## Lex Recall Reverse (Median)
p = 0.60 or more, doesn't matter where the median is in. 
```{r}
lex_fw_hl <- eyeperf %>% 
  select(lex_reversed, chest_reversed) 
lex_mean = median(lex_fw_hl$lex_reversed, na.rm = TRUE)

lex_fw_hl <- lex_fw_hl %>%
  mutate(group = case_when(
    lex_reversed > lex_mean ~ "high",
    lex_reversed <= lex_mean ~ "low")) %>%
  mutate(group = factor(group))

summary(lex_fw_hl)

lex_fw_high = filter(lex_fw_hl, group == "high") %>% select(chest_reversed)
lex_fw_low = filter(lex_fw_hl, group == "low") %>% select(chest_reversed)

lexfw_high = unlist(lex_fw_high$chest_reversed)
lexfw_low = unlist(lex_fw_low$chest_reversed)
t.test(lexfw_high, lexfw_low)
```

# Top v Bottom 25% Performers

Let's try this again but with top and bottom 25% people. 

"It makes total sense to me why that [high v. low] approach doesn't work, because many who are right around the median don't help address what happens when you have GOOD comprehension (i.e. ABOVE average) vs POOR, especially if you have a lot who straddle the median.   SO you need to move toward the ends."

## Gist Reversed

Well, that's not going to work becuase we just have three possible scores - 0, 0.5, and 1. I'll try that as three groups! And do an ANOVA. 
```{r}
gist_hl <- eyeperf %>% 
  select(gist_reversed, chest_reversed) %>%
  mutate(group = case_when(
    gist_reversed == 0 ~ "low",
    gist_reversed == 0.5 ~ "middle",
    gist_reversed == 1 ~ "high"
  )) %>%
  mutate(group = factor(group)) %>%
  select(-gist_reversed)

id = seq(1,nrow(gist_hl))
gist_hl = cbind(gist_hl, id)
  
  

g1 <- aov_ez("id", "chest_reversed", gist_hl, between = "group")
g1
```

## Lex Recall Forward 
p = 0.656 (doesn't matter if inclusive or exclusive)
```{r}
lex_fw_hl <- eyeperf %>% 
  select(lex_forward, chest_forward) 
lex_quartiles = quantile(lex_fw_hl$lex_forward, probs = c(0.25, 0.75))

lex_fw_hl <- lex_fw_hl %>%
  mutate(group = case_when(
    lex_forward >= lex_quartiles[2] ~ "high",
    lex_forward <= lex_quartiles[1] ~ "low")) %>%
  mutate(group = factor(group)) %>%
  filter(!is.na(group))

summary(lex_fw_hl)

lex_fw_high = filter(lex_fw_hl, group == "high") %>% select(chest_forward)
lex_fw_low = filter(lex_fw_hl, group == "low") %>% select(chest_forward)

lexfw_high = unlist(lex_fw_high$chest_forward)
lexfw_low = unlist(lex_fw_low$chest_forward)
t.test(lexfw_high, lexfw_low)
```

## Lex Recall Reverse 
p = 0.656 (doesn't matter if inclusive or exclusive)
```{r}
lex_fw_hl <- eyeperf %>% 
  select(lex_reversed, chest_reversed) %>%
  filter(!is.na(lex_reversed))
lex_quartiles = quantile(lex_fw_hl$lex_reversed, probs = c(0.25, 0.75))

lex_fw_hl <- lex_fw_hl %>%
  mutate(group = case_when(
    lex_reversed >= lex_quartiles[2] ~ "high",
    lex_reversed <= lex_quartiles[1] ~ "low")) %>%
  mutate(group = factor(group)) %>%
  filter(!is.na(group))

summary(lex_fw_hl)

lex_fw_high = filter(lex_fw_hl, group == "high") %>% select(chest_reversed)
lex_fw_low = filter(lex_fw_hl, group == "low") %>% select(chest_reversed)

lexfw_high = unlist(lex_fw_high$chest_reversed)
lexfw_low = unlist(lex_fw_low$chest_reversed)
t.test(lexfw_high, lexfw_low)
```

## Scatterplots
```{r}
library(scales)
ggplot(eyeperf, aes(x = gist_reversed, y = facechest_reversed)) + 
  geom_point() + geom_smooth(method = "lm") +
  scale_x_continuous(labels = percent) +
  scale_y_continuous(limits = c(-1,1)) + ggtitle("Gist Rv")
ggplot(eyeperf, aes(x = lex_forward, y = facechest_forward)) + 
  geom_point() + geom_smooth(method = "lm") +
  scale_x_continuous(labels = percent) +
  scale_y_continuous(limits = c(-1,1)) + ggtitle("Lex Fw")
ggplot(eyeperf, aes(x = lex_reversed, y = facechest_reversed)) + 
  geom_point() + geom_smooth(method = "lm") +
  scale_x_continuous(labels = percent) +
  scale_y_continuous(limits = c(-1,1)) + ggtitle("Lex Rv")
```

# Repeated Measures Correlations
OKAY! I found [this article](https://www.frontiersin.org/articles/10.3389/fpsyg.2017.00456/full#h11) which discusses correlations with repeated measures, and it involves the `rmcorr` package. Let's give that a spin. 

```{r}
library(rmcorr)
data <- fulldata

data <- data %>% 
  select(participant, direction, maingroup, aoasl, acc, gist, facechest, eyes, mouth, neck, upperchest) %>%
  arrange(participant, direction)

ggplot(data, aes(x = acc, y = facechest)) + geom_point() + geom_smooth(method = "lm") + xlab("lex recall (all)") + scale_x_continuous(labels = percent) + ylab("facechest (all)")
ggplot(data, aes(x = acc, y = facechest, color = maingroup)) + geom_point() + geom_smooth(method = "lm") + xlab("lex recall") + scale_x_continuous(labels = percent) + ylab("facechest") + facet_grid(direction ~ maingroup) + guides(color = FALSE)

data_fw <- data %>% filter(direction == "forward")
data_rv <- data %>% filter(direction == "reversed")

fw <- rmcorr(participant, facechest, acc, data_fw)
rv <- rmcorr(participant, facechest, acc, data_rv)
all <- rmcorr(participant, facechest, acc, data)

plot(fw, data_fw)
plot(rv, data_rv)
plot(all, data)

```

# Regular Correlation by Groups
## Deaf Early
```{r rows.print = 15}
corstarsl2 <- function(x){ 
require(Hmisc) 
x <- as.matrix(x) 
R <- Hmisc::rcorr(x)$r 
p <- Hmisc::rcorr(x)$P 
## define notions for significance levels; spacing is important.
mystars <- ifelse(p < .001, "***", ifelse(p < .01, "** ", ifelse(p < .05, "* ", " ")))
## trunctuate the matrix that holds the correlations to two decimal
R <- format(round(cbind(rep(-1.11, ncol(x)), R), 2))[,-1] 
## build a new matrix that includes the correlations with their apropriate stars 
Rnew <- matrix(paste(R, mystars, sep=""), ncol=ncol(x)) 
diag(Rnew) <- paste(diag(R), " ", sep="") 
rownames(Rnew) <- colnames(x) 
colnames(Rnew) <- paste(colnames(x), "", sep="") 
## remove upper triangle
Rnew <- as.matrix(Rnew)
Rnew[upper.tri(Rnew, diag = TRUE)] <- ""
Rnew <- as.data.frame(Rnew) 
## remove last column and return the matrix (which is now a data frame)
Rnew <- cbind(Rnew[2:3])
return(Rnew) 
}
eyeperf %>% filter(maingroup == "DeafEarly") %>% select(lex_forward, 
                                                       lex_reversed, 
                                                       facechest_forward, 
                                                       facechest_reversed) %>%
  corstarsl(.)
eyeperf1 <- eyeperf %>% filter(maingroup == "DeafEarly") %>% 
  select(lex_forward, facechest_forward) %>%
  rename(lex = lex_forward,
         facechest = facechest_forward)

eyeperf2 <- eyeperf %>% filter(maingroup == "DeafEarly") %>% 
  select(lex_reversed, facechest_reversed) %>%
  rename(lex = lex_reversed,
         facechest = facechest_reversed)

rbind(eyeperf1, eyeperf2) %>% corstarsl(.)
  
```

```{r rows.print = 15}
eyeperf %>% filter(maingroup == "DeafLate") %>% select(lex_forward, 
                                                       lex_reversed, 
                                                       facechest_forward, 
                                                       facechest_reversed) %>%
  corstarsl(.)
eyeperf1 <- eyeperf %>% filter(maingroup == "DeafLate") %>% 
  select(lex_forward, facechest_forward) %>%
  rename(lex = lex_forward,
         facechest = facechest_forward)

eyeperf2 <- eyeperf %>% filter(maingroup == "DeafLate") %>% 
  select(lex_reversed, facechest_reversed) %>%
  rename(lex = lex_reversed,
         facechest = facechest_reversed)

rbind(eyeperf1, eyeperf2) %>% corstarsl(.)

```

```{r rows.print = 15}
eyeperf %>% filter(maingroup == "HearingLate") %>% select(lex_forward, 
                                                       lex_reversed, 
                                                       facechest_forward, 
                                                       facechest_reversed) %>%
  corstarsl(.)

eyeperf1 <- eyeperf %>% filter(maingroup == "HearingLate") %>% 
  select(lex_forward, facechest_forward) %>%
  rename(lex = lex_forward,
         facechest = facechest_forward)

eyeperf2 <- eyeperf %>% filter(maingroup == "HearingLate") %>% 
  select(lex_reversed, facechest_reversed) %>%
  rename(lex = lex_reversed,
         facechest = facechest_reversed)

rbind(eyeperf1, eyeperf2) %>% corstarsl(.)
```

```{r rows.print = 15}
eyeperf %>% filter(maingroup == "HearingNovice") %>% select(lex_forward, 
                                                       lex_reversed, 
                                                       facechest_forward, 
                                                       facechest_reversed) %>%
  corstarsl(.)

eyeperf1 <- eyeperf %>% filter(maingroup == "HearingNovice") %>% 
  select(lex_forward, facechest_forward) %>%
  rename(lex = lex_forward,
         facechest = facechest_forward)

eyeperf2 <- eyeperf %>% filter(maingroup == "HearingNovice") %>% 
  select(lex_reversed, facechest_reversed) %>%
  rename(lex = lex_reversed,
         facechest = facechest_reversed)

rbind(eyeperf1, eyeperf2) %>% corstarsl(.)

```

```{r}
newfcr <- fulldata %>%
  filter(eye_exclude == FALSE) %>%
  select(participant, maingroup, direction, acc, forehead, eyes, mouth, neck, upperchest, midchest, lowerchest, belly, facechest) %>% 
  rowwise() %>%
  mutate(face = sum(forehead, eyes, mouth, na.rm = TRUE),
         chest = sum(upperchest, midchest, lowerchest, na.rm = TRUE),
         fcr2 = (face - chest)/(face+chest)) %>%
  select(maingroup, direction, acc, facechest, fcr2) %>%
  arrange(facechest)

newfcr %>% group_by(maingroup) %>%
  summarise(fcr_mean = mean(facechest),
            fcr_sd = sd(facechest),
            fcr2_mean = mean(fcr2),
            fcr2_sd = sd(fcr2))

ggplot(newfcr, aes(x = acc, y = fcr2, color = maingroup)) + geom_point() +
  facet_grid(direction ~ maingroup) + geom_smooth(method = "lm")
```

