---
title: "Study 1 Adult Data Analysis"
author: "Adam Stone, PhD"
date: "August 31, 2017"
output:
  # github_document:
  #   toc: yes
  #   toc_depth: 2
  html_notebook:
    code_folding: hide
    theme: united
    toc: yes
    toc_depth: 2
    toc_float: yes
---

## Importing and Reshaping Data
Here we're going to import the data, remove dropped participants, and reshape the data so story and direction are grouping variables (and the dataset will be more tall than wide). Let's see all our data first (scroll horizontally).

```{r, echo=FALSE, message=FALSE, warning=FALSE}

# Import packages we'll need.
library(tidyverse)
library(stringr)
library(lme4)
library(prettydoc)
library(broom)
library(knitr)
library(xtable)
library(kableExtra)
options(knitr.table.format = "html") 
# Import the data
data <- read_csv('finaladultdata.csv', col_types=
                   cols(
                     id = col_integer(),
                     participant = col_character(),
                     hearing = col_character(),
                     videogroup = col_character(),
                     aoagroup = col_character(),
                     languagegroup = col_character(),
                     maingroup = col_character(),
                     selfrate = col_double(),
                     age = col_double(),
                     signyrs = col_double(),
                     aoasl = col_integer(),
                     acc.fw1 = col_double(),
                     acc.rv2 = col_double(),
                     acc.fw3 = col_double(),
                     acc.rv4 = col_double(),
                     forehead.fw1 = col_double(),
                     forehead.fw3 = col_double(),
                     forehead.rv2 = col_double(),
                     forehead.rv4 = col_double(),
                     eyes.fw1 = col_double(),
                     eyes.fw3 = col_double(),
                     eyes.rv2 = col_double(),
                     eyes.rv4 = col_double(),
                     mouth.fw1 = col_double(),
                     mouth.fw3 = col_double(),
                     mouth.rv2 = col_double(),
                     mouth.rv4 = col_double(),
                     chin.fw1 = col_double(),
                     chin.fw3 = col_double(),
                     chin.rv2 = col_double(),
                     chin.rv4 = col_double(),
                     upperchest.fw1 = col_double(),
                     upperchest.fw3 = col_double(),
                     upperchest.rv2 = col_double(),
                     upperchest.rv4 = col_double(),
                     midchest.fw1 = col_double(),
                     midchest.fw3 = col_double(),
                     midchest.rv2 = col_double(),
                     midchest.rv4 = col_double(),
                     lowerchest.fw1 = col_double(),
                     lowerchest.fw3 = col_double(),
                     lowerchest.rv2 = col_double(),
                     lowerchest.rv4 = col_double(),
                     belly.fw1 = col_double(),
                     belly.fw3 = col_double(),
                     belly.rv2 = col_double(),
                     belly.rv4 = col_double(),
                     left.fw1 = col_double(),
                     left.fw3 = col_double(),
                     left.rv2 = col_double(),
                     left.rv4 = col_double(),
                     right.fw1 = col_double(),
                     right.fw3 = col_double(),
                     right.rv2 = col_double(),
                     right.rv4 = col_double()
                     )
)
kable(data) %>% kable_styling(bootstrap_options = c("striped", "hover","condensed","responsive"), font_size=7)
```

Remove dropped participants (they have no main group name, or no data). These are who we dropped. 

```{r, message=FALSE, warning=FALSE}
dropped <- filter(data, is.na(maingroup)==TRUE)
manual <- filter(data,participant=="Lucinda" | participant=="Joe")
dropped <- rbind(dropped,manual) %>% arrange(id)
data <- data %>%
  filter(is.na(maingroup)==FALSE) %>% 
  filter(participant!="Lucinda" & participant!="Joe")
kable(dropped) %>% kable_styling(bootstrap_options = c("striped", "hover","condensed"))
```

Now we'll reshape the data. Based on Rain's UNM talk, this is what Group 1 & 2 saw:

| No. | Group1 | Group2 |
|:------|:-----|:------|
| 1     | Red Riding Hood Fwd    | Goldilocks Fwd   |
| 2     | King Midas Rev    | Cinderella Rev   |
| 3    | Cinderella Fwd    | King Mias Fwd   |
| 4    | Goldilocks Rev    | Red Riding Hood Rev   |

Let's add that information to our data when we reshape it. 

```{r, message=FALSE, warning=FALSE}
# I tried writing a function to do this using column names as arguments. 
# But after hours of googling I couldn't figure it out! So just copying/pasting code here.
data.acc <- data %>%
  select(id,acc.fw1:acc.rv4) %>%
  gather(video,acc,acc.fw1:acc.rv4, factor_key=TRUE) %>%
  mutate(video = str_sub(video,-3,-1))
data.forehead <- data %>%
  select(id,forehead.fw1:forehead.rv4) %>%
  gather(video,forehead,forehead.fw1:forehead.rv4) %>%
  mutate(video = str_sub(video,-3,-1))
data.eyes <- data %>%
  select(id,eyes.fw1:eyes.rv4) %>%
  gather(video,eyes,eyes.fw1:eyes.rv4) %>%
  mutate(video = str_sub(video,-3,-1))
data.mouth <- data %>%
  select(id,mouth.fw1:mouth.rv4) %>%
  gather(video,mouth,mouth.fw1:mouth.rv4) %>%
  mutate(video = str_sub(video,-3,-1))
data.chin <- data %>%
  select(id,chin.fw1:chin.rv4) %>%
  gather(video,chin,chin.fw1:chin.rv4) %>%
  mutate(video = str_sub(video,-3,-1))
data.upperchest <- data %>%
  select(id,upperchest.fw1:upperchest.rv4) %>%
  gather(video,upperchest,upperchest.fw1:upperchest.rv4) %>%
  mutate(video = str_sub(video,-3,-1))
data.midchest <- data %>%
  select(id,midchest.fw1:midchest.rv4) %>%
  gather(video,midchest,midchest.fw1:midchest.rv4) %>%
  mutate(video = str_sub(video,-3,-1))
data.lowerchest <- data %>%
  select(id,lowerchest.fw1:lowerchest.rv4) %>%
  gather(video,lowerchest,lowerchest.fw1:lowerchest.rv4) %>%
  mutate(video = str_sub(video,-3,-1))
data.belly <- data %>%
  select(id,belly.fw1:belly.rv4) %>%
  gather(video,belly,belly.fw1:belly.rv4) %>%
  mutate(video = str_sub(video,-3,-1))
data.left <- data %>%
  select(id,left.fw1:left.rv4) %>%
  gather(video,left,left.fw1:left.rv4) %>%
  mutate(video = str_sub(video,-3,-1))
data.right <- data %>%
  select(id,right.fw1:right.rv4) %>%
  gather(video,right,right.fw1:right.rv4) %>%
  mutate(video = str_sub(video,-3,-1))
data.header <- data[1:11]

# Join them all back together. inner_join is smart, it'll join by id AND video.
newdata <- data.header %>%
  inner_join(data.acc,by="id") %>%
  inner_join(data.forehead) %>%
  inner_join(data.eyes) %>%
  inner_join(data.mouth) %>%
  inner_join(data.chin) %>%
  inner_join(data.upperchest) %>%
  inner_join(data.midchest) %>%
  inner_join(data.lowerchest) %>%
  inner_join(data.belly) %>%
  inner_join(data.left) %>%
  inner_join(data.right)

# Let's change the names
oldata <- data
data <- newdata

# Add story and direction variables and split data into videogroups
data$story <- NA
data$direction <- NA
group1 <- filter(data,videogroup=="Group 1")
group2 <- filter(data,videogroup=="Group 2")

# Now define levels for story and direction based on that table above
group1 <- mutate(group1,story = ifelse(video == "fw1","RedRidingHood",
                                       ifelse(video == "rv2","KingMidas",
                                              ifelse(video =="fw3","Cinderella","Goldilocks"))))
group1 <- mutate(group1,direction = ifelse(video == "fw1","forward",
                                       ifelse(video == "rv2","reversed",
                                              ifelse(video == "fw3","forward","reversed"))))
group2 <- mutate(group2,story = ifelse(video == "fw1","Goldilocks",
                                       ifelse(video == "rv2","Cinderella",
                                              ifelse(video=="fw3","KingMidas","RedRidingHood"))))
group2 <- mutate(group2,direction = ifelse(video == "fw1","forward",
                                           ifelse(video == "rv2","reversed",
                                                  ifelse(video == "fw3","forward","reversed"))))

# Join groups back together and view
data <- rbind(group1,group2)
data <- arrange(data,id,video)

# Convert some columns to factors
data <- data %>%
  mutate(hearing = as.factor(hearing)) %>%
  mutate(videogroup = as.factor(videogroup)) %>%
  mutate(aoagroup = as.factor(aoagroup)) %>%
  mutate(languagegroup = as.factor(languagegroup)) %>%
  mutate(maingroup = as.factor(maingroup)) %>%
  mutate(video = as.factor(video)) %>%
  mutate(story = as.factor(story)) %>%
  mutate(direction = as.factor(direction)) %>%
  select(id,participant,hearing,videogroup,aoagroup,languagegroup,maingroup,video,story,
         direction,age,selfrate,signyrs,aoasl,acc,forehead,eyes,mouth,chin,upperchest,
         midchest,lowerchest,belly,left,right)

```
Here's the final, "cleaned-up" dataset that we're going to use for all further analysis. Just the first 10 rows. Notice that there are now story and direction columns, and AOI-only columns, and each participant has 4 observation rows (one observation per story). So, more rows, less columns.  
```{r}
kable(head(data, n=10)) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

## Participant Demographics
Now we can easily get group means.
```{r}
groupmeans <- data %>%
  group_by(maingroup) %>%
  summarize(n = n()/4,
            age = mean(age),
            selfrate = mean(selfrate),
            signyrs = mean(signyrs),
            aoasl = mean(aoasl))
kable(groupmeans)
```

## Accuracy Data Analysis
And accuracy boxplots and error bar charts for forward vs. backward stories. 
```{r}
# Summarizing means and SDs
accdata <- data %>%
  group_by(maingroup,direction) %>%
  summarize(acc.mean = mean(acc, na.rm=TRUE),
            acc.sd = sd(acc, na.rm=TRUE))

#Boxplot
ggplot(data,aes(maingroup,acc,fill=direction)) + 
  geom_boxplot() +
  scale_y_continuous(limits=c(0,1)) +
  theme(axis.text.x=element_text(angle=45,hjust=1))
```
```{r}
# Error bar chart
ggplot(accdata,aes(maingroup,acc.mean,color=direction)) + 
  geom_point(position=position_dodge(0.5)) +
  geom_errorbar(aes(ymin=acc.mean-acc.sd,ymax=acc.mean+acc.sd),width=0.1,position=position_dodge(0.5)) +
  scale_y_continuous(limits=c(0,1)) +
  theme(axis.text.x=element_text(angle=45,hjust=1))

```

Let's test for statistical significance. A simple ANOVA tell us there is a main effect of group and direction, but no interactions. Tukey's HSD posthoc tells us that Hearing Novice ASL is significantly different from Deaf Native.  
```{r}
# Let's set Native Deaf as the reference level to compare all other
data$maingroup <- relevel(data$maingroup, ref="NativeDeaf")
# Run the ANOVA
acc.anova <- aov(data=data,acc ~ maingroup*direction)
kable(xtable(acc.anova))
# Run the posthoc on main group
TukeyHSD(acc.anova,'maingroup',conf.level = 0.95) 
```

Group coefficients are here. Remember our reference level ("control") is Native Deaf. Their forward accuracy is 85% with a reversal effect of -9%.  ASL, Forward. Mean accuracy for that is 84%, and its reversal effect is -14%. All the other values are to be added to these coefficients. Hearing Novice ASL's forward accuracy was 78%, with a reversal effect of -6%! 
```{r}
#Coefficients
acc.anova$coefficients
```

In summary, ANOVA tells us there are main effects of group and direction, no interactions. All of which is a good thing. But I'm curious if there's any item-level effects we should be watching out for. Because there are 4 different stories. Let's plot those out.

```{r}
# Run summary stats grouped by story, too
accdata2 <- data %>%
  group_by(maingroup,story,direction) %>%
  summarize(acc.mean = mean(acc, na.rm=TRUE),
            acc.sd = sd(acc, na.rm=TRUE))
# Boxplot
# ggplot(data,aes(maingroup,acc,fill=direction)) + 
#   geom_boxplot() +
#   scale_y_continuous(limits=c(0,1)) +
#   facet_wrap("story")

# Error bar chart
ggplot(accdata2,aes(maingroup,acc.mean,color=direction)) + 
  geom_point(position=position_dodge(0.5)) +
  geom_errorbar(aes(ymin=acc.mean-acc.sd,ymax=acc.mean+acc.sd),width=0.1,position=position_dodge(0.5)) +
  scale_y_continuous(limits=c(0,1)) +
  facet_wrap("story") + 
  theme(axis.text.x=element_text(angle=45,hjust=1))
```

Boy! Seems King Midas had a strong reversal effect, while Red Riding Hood had a weak reversal effect. Maybe we should put those in as random effects variables in a mixed model, along with participants too. With mixed models, you define predictor variables (what we're interested in; aka, fixed effects) and grouping (what we're not interested in, aka, random effects). This is overkill for simple accuracy data but this will help set us up for eye tracking analysis and **importantly reviewers may ask us about item-level effects given we have just 4 stories.**

So here, we have fixed effects of group and direction, and random effects of story and id, with varying slopes for direction.

The output tells us now that Hearing Novice ASL is significantly different from Deaf Native ASL, that we still have a main effect of direction, with a  significant interaction of direction and Hearing Late ASL. We probably don't need all this in the paper, though, the ANOVA will do and accuracy is not our main point. 
```{r}
acc.lm <- lmer(data=data, acc ~ maingroup*direction + (direction|id) + (1|story))
summary(acc.lm)
```


## Eye Data 
To come!